### 2023-08-30 2_scripts/2.0 AB Outpatient Lab Data.qmd
Upload APL data cleaning code

---
title: "2.0 AB Outpatient Lab Data Encryption (SHA-2)"
author: J. Chen
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
# Load Paks
library(dplyr)
library(duawranglr)    # use this to encrypt the ids
```

```{r}
## Import Dataset
library(readxl)
RFD4682_COVID<-read_excel("~/APL/RFD4682_COVID_Serology_Data_for release June 2023.xlsx", 
    sheet = "RFD4682_Full_Data", col_types = c("text", 
        "text", "text", "text", "text", 
        "date", "text", "text", "date", "date", 
        "text", "text", "numeric", "text", 
        "text", "text", "text", "text", "date", 
        "numeric", "text", "text", 
        "date", "numeric", "numeric"))

table(RFD4682_COVID$NAAT_STATUS, useNA = 'ifany')
save(RFD4682_COVID, file = "~/APL/RFD4682.RData")

## The IDs are not encrypted, need to do it by ourselves

## Add missing value indicator for PERSON_ID and ULI_ID
RFD4682_COVID$id_per_miss<-is.na(RFD4682_COVID$PERSON_ID)
RFD4682_COVID$id_uli_miss<-is.na(RFD4682_COVID$ULI_ID)

# Use deid_dua() from {duawranglr} package to do the encryption
RFD4682_e <- deid_dua(RFD4682_COVID, 
                      id_col = 'PERSON_ID',
                      new_id_name = 'PERSON_IDe',
                      write_crosswalk = TRUE, id_length = 12,
                      crosswalk_filename='crosswalk_id_per.csv')

RFD4682_e <- deid_dua(RFD4682_e, 
                      id_col = 'ULI_ID',
                      new_id_name = 'ULI_IDe',
                      write_crosswalk = TRUE, id_length = 15)

# SHA-2 algorithm also process missing values in those two ids
# NAs in each id will have the same encrypted value, need to replace them with NA
RFD4682_e$PERSON_IDe[RFD4682_e$id_per_miss==TRUE]<-NA
RFD4682_e$ULI_IDe[RFD4682_e$id_uli_miss==TRUE]<-NA
RFD4682_e<-select(RFD4682_e, !c(id_per_miss, id_uli_miss))

## Take a look at the crosswalk file
corsswalk_uli<-read.csv(file = '~/APL/id_crosswalk_id_uli.csv')
rm(corsswalk_uli)
# Alton is the only person who keeps the crosswalks

## Save the Dataset
save(RFD4682_e, file = "~/APL/RFD4682e.RData")
```

## Check the unique individual in the original dataset

```{r}
## ULI_ID: Unique Lifetime Identifier (ULI), numerical values of Alberta PHN.
length(unique(RFD4682_e$ULI_IDe))     # Unique = 180821-1 (1 is NA)
table(is.na(RFD4682_e$ULI_IDe))       # NA = 11298

## PERSON_ID: assigned from the originating LIS
length(unique(RFD4682_e$PERSON_IDe))     # Unique = 189864-1
table(is.na(RFD4682_e$PERSON_IDe))       # NA = 2973
```

## Matt's Data
```{r}

```