### 2023-08-30 2_scripts/2.1 APL Data Quality Check.qmd

---
title: "2.1 APL Data Quality Check with additonal cleaning"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

## Load the APL data with encrypted ID variables
```{r}
load(file = "~/APL/RFD4682e.RData")
```

## Load paks
```{r}
library(dplyr)
```

```{r}
## Create subset for those NAAT Positive Cases
table(RFD4682_e$NAAT_STATUS, useNA = 'ifany')
RFD4682_NAAT<-RFD4682_e[RFD4682_e$NAAT_STATUS %in% c('Positive'),]

# Save the subset

# Load Data file

```

## The distribution of time from NAAT to Sero-assay

```{r}
hist(RFD4682_e$DAYS_SINCE_NAAT[RFD4682_e$NAAT_STATUS=="Positive"], breaks = 18,
     main="Time form NAAT to Sero-assay (Days)",
     xlab = "Days")

## The time variable is good to go
```

## Records are not patients, how many individual has been included?
```{r}
## ULI_ID: Unique Lifetime Identifier (ULI), numerical values of Alberta PHN.
length(unique(RFD4682_e$ULI_IDe))     # Unique = 180821
table(is.na(RFD4682_e$ULI_IDe))       # NA = 11298

## PERSON_ID: assigned from the originating LIS
length(unique(RFD4682_e$PERSON_IDe))     # Unique = 189864
table(is.na(RFD4682_e$PERSON_IDe))       # NA = 2973

## SERO_ID: Unique Specimen ID + Test month, unique APL dataset
length(unique(RFD4682_e$SERO_ID)) # N=214,772
table(is.na(RFD4682_e$SERO_ID))   # No missing

## Checking duplicated records by SERO_ID
tab_SERO_ID<-table(RFD4682_e$SERO_ID)
View(tab_SERO_ID)

str(tab_SERO_ID)
tab_SERO_ID.dat<-as.data.frame(tab_SERO_ID)
tab_SERO_ID.dat[tab_SERO_ID.dat$Freq==2,]
```
## Why unique PERSON_ID and ULI_ID are not matching with each other?
```{r}
## Check the discrepancies between those two

## Freq tab for ULI_ID
tab_ULI_IDe<-table(RFD4682_e$ULI_IDe)
tab_ULI_IDe.dat<-as.data.frame(tab_ULI_IDe)
colnames(tab_ULI_IDe.dat)<-c('ULI_IDe','Freq_ULI')

# tab_ULI_ID.dat.2<-tab_ULI_ID.dat[tab_ULI_ID.dat$Freq>1,]

## Freq tab for RERSON_ID
tab_PERSON_IDe<-table(RFD4682_e$PERSON_IDe)
tab_PERSON_IDe.dat<-as.data.frame(tab_PERSON_IDe)
colnames(tab_PERSON_IDe.dat)<-c('PERSON_IDe','Freq_PER')

## Add freq var to the subset
RFD4682_e$order_ID<-1:214780
RFD4682_3<-select(RFD4682_e, order_ID, SERO_ID, ULI_IDe, PERSON_IDe)
RFD4682_3<-merge(RFD4682_3, tab_ULI_IDe.dat, by='ULI_IDe',
      all.x = T, all.y = F)
RFD4682_3<-merge(RFD4682_3, tab_PERSON_IDe.dat, by='PERSON_IDe',
      all.x = T, all.y = F)

## Find Discrepancies by compare the frequencies

## NA is messy with logic, replace it with 0
RFD4682_3$Freq_ULI[is.na(RFD4682_3$Freq_ULI)==T]<-0
RFD4682_3$Freq_PER[is.na(RFD4682_3$Freq_PER)==T]<-0
tab.idfreq<-table(RFD4682_3$Freq_ULI, RFD4682_3$Freq_PER, useNA = 'ifany')
tab.idfreq
dat.idfreq<-as.data.frame(tab.idfreq)

# Scenario 0: Frequencies of ULI and Person_ID are both missing (equal too)
RFD4682_3$ID_clean<-rep(NA, 214780)
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI==0 &
                   RFD4682_3$Freq_PER==0  ]<-0

# Scenario 1: Frequencies of ULI and Person_ID are equal
length(unique(RFD4682_3$ULI_IDe[(RFD4682_3$Freq_ULI==RFD4682_3$Freq_PER) &
                                 RFD4682_3$Freq_ULI>0                    &
                                 RFD4682_3$Freq_PER>0])
) # N=179,460 This is the majority of cases in the dataset, which is good
# This number matches with the numbers from the confusion tab

RFD4682_3$ID_clean[(RFD4682_3$Freq_ULI==RFD4682_3$Freq_PER) &
                                 RFD4682_3$Freq_ULI>0       &
                                 RFD4682_3$Freq_PER>0]<-1

# Scenario 2: only ULI, Person_ID are missing
# Number estimated using confusion tab: 107, 108 records
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI>0 &
                   RFD4682_3$Freq_PER==0  ]<-2

# Scenario 3, ULI missing but they have PERSON_ID
# Number estimated using confusion tab: 7904, 8433 records
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI==0 &
                   RFD4682_3$Freq_PER>0  ]<-3

# Scenario 4, one PERSON_ID matches with multiple ULI_IDs
# Number estimated using confusion tab: 6, 10 records
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI==1 &
                   RFD4682_3$Freq_PER==2  ]<-4 # this would only fix 6 records

# Fix the rest 4 manually
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="9a4299b85800" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="aa6526b42bfe" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="e054534f4352" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="fcb0ca36fbb4" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4

RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="9a4299b85800" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"3c32a58470ee228"
RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="aa6526b42bfe" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"f21a83130a09b98"
RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="e054534f4352" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"c80b7fbde166521"
RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="fcb0ca36fbb4" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"08bc00a9a5e8720"

# The rest are scenario 5
# N records = 2775, number matches those on confusion cross-tab
RFD4682_3$ID_clean[is.na(RFD4682_3$ID_clean)==TRUE]<-5

# Rename the variable as "scenario"
RFD4682_3<-rename(RFD4682_3, Scenario = ID_clean)
table(RFD4682_3$Scenario, useNA = 'ifany')

# Apply approperiate individual ids to each records, according to its scenario
RFD4682_3$clean_IDe<-rep(NA, 214780)
RFD4682_3$clean_IDe<-if_else(RFD4682_3$Scenario %in% c(3), RFD4682_3$PERSON_IDe, RFD4682_3$ULI_IDe)
```

## Scenario Examples Showcase
```{r}
# Example with each Scenarios
RFD4682_3[RFD4682_3$Freq_ULI==6 & RFD4682_3$Freq_PER==5 ,]     # Scenario 5
RFD4682_3[RFD4682_3$Freq_ULI==15 & RFD4682_3$Freq_PER==15 ,]   # Scenario 1
RFD4682_3[RFD4682_3$Freq_ULI==2 & RFD4682_3$Freq_PER==0 ,]     # Scenario 2
RFD4682_3[RFD4682_3$Freq_ULI==0 & RFD4682_3$Freq_PER==5 ,]     # Scenario 3
RFD4682_3[RFD4682_3$Freq_ULI==1 & RFD4682_3$Freq_PER==2 ,]     # Scenario 4
RFD4682_3[RFD4682_3$Freq_ULI==0 & RFD4682_3$Freq_PER==0 ,]     # Scenario 0
```

## Add clean_IDe and Scenario back to dataset (encrypted)
```{r}
RFD4682_e<-merge(RFD4682_e, select(RFD4682_3, order_ID, Scenario, clean_IDe),
                      by = 'order_ID', all = TRUE)
```

#### ULI ID and PERSON_ID, in NAAT+ subset
```{r}
## ULI_ID: Unique Lifetime Identifier (ULI), numerical values of Alberta PHN.
length(unique(RFD4682_NAAT$ULI_ID))     # N = 5933
table(is.na(RFD4682_NAAT$ULI_ID))       # NA = 0

## PERSON_ID: assigned from the originating LIS
length(unique(RFD4682_NAAT$PERSON_ID)) # N = 5940
table(is.na(RFD4682_NAAT$PERSON_ID))   # NA = 0

## Check the discrepancies between those two
tab_ULI_ID<-table(RFD4682_NAAT$ULI_ID)
tab_ULI_ID.dat<-as.data.frame(tab_ULI_ID)
colnames(tab_ULI_ID.dat)<-c('ULI_ID','Freq_ULI')
tab_ULI_ID.dat.2<-tab_ULI_ID.dat[tab_ULI_ID.dat$Freq>1,]

tab_PERSON_ID<-table(RFD4682_NAAT$PERSON_ID)
tab_PERSON_ID.dat<-as.data.frame(tab_PERSON_ID)
colnames(tab_PERSON_ID.dat)<-c('PERSON_ID','Freq_PER')
tab_PERSON_ID.dat.2<-tab_PERSON_ID.dat[tab_PERSON_ID.dat$Freq>1,]

## Add freq var to 
RFD4682_NAAT_3<-select(RFD4682_NAAT, SERO_ID, ULI_ID, PERSON_ID)
RFD4682_NAAT_3<-merge(RFD4682_NAAT_3, tab_ULI_ID.dat.2, by='ULI_ID',
      all.x = T, all.y = F)
RFD4682_NAAT_3<-merge(RFD4682_NAAT_3, tab_PERSON_ID.dat.2, by='PERSON_ID',
      all.x = T, all.y = F)

RFD4682_NAAT_4<-RFD4682_NAAT_3[is.na(RFD4682_NAAT_3$Freq_ULI)==F |
               is.na(RFD4682_NAAT_3$Freq_PER)==F,]

RFD4682_NAAT_4$freq_minus<-RFD4682_NAAT_4$Freq_ULI-RFD4682_NAAT_4$Freq_PER

print(RFD4682_NAAT_4[RFD4682_NAAT_4$freq_minus>0 |
               is.na(RFD4682_NAAT_4$freq_minus)==TRUE,])

## Save the dataset
save(RFD4682_e, file = '~/APL/RFD4682e.RData')

## Conclusion:
## Use ULI_ID, in NAAT+ subset, seven people have 2 diff PERSON_IDs
## Unique ULI_ID: N=5940
```

```{r}
length(unique(RFD4682_e$clean_IDe))
table(is.na(unique(RFD4682_e$clean_IDe)), useNA = 'ifany')

# Unique Individual in the dataset: 188,722 (beware of missing values)
```

## Check N=2865 cases with missing in both ULI and Person_ID
```{r}
RFD4682_e[RFD4682_e$Scenario==0,]
## Surprisingly, most of the records in this group has some useful variables,
## such as: age, gender and Anti_RBD test results

# Age
table(RFD4682_e$AGE_AT_COLLECTION[RFD4682_e$Scenario==0], useNA = 'ifany') # 0.03%
# Gender
table(RFD4682_e$GENDER[RFD4682_e$Scenario==0], useNA = 'ifany') # 0.03%
# N-Ig
table(RFD4682_e[RFD4682_e$Scenario==0,]$`N-IgG_RESULT`, useNA = 'ifany') # Can't show all
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$`N-IgG_RESULT`), useNA = 'ifany') # 92.95%
table(is.na(RFD4682_e$`N-IgG_RESULT`), useNA = 'ifany') # 92.95%

# Ig Anti-RBD
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$`RBD-IgGII_RESULT`), useNA = 'ifany')
# no missing, good

# FSA
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$PAT_FSA), useNA = 'ifany') # 9.49%

# Vaccination
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$VACCINATION_STATUS), useNA = 'ifany')
# No miss, good
table(RFD4682_e[RFD4682_e$Scenario==0,]$VACCINATION_STATUS, useNA = 'ifany')
table(RFD4682_e$VACCINATION_STATUS, useNA = 'ifany')

# NAAT Status
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$NAAT_STATUS), useNA = 'ifany') # 92.98%
table(is.na(RFD4682_e$NAAT_STATUS), useNA = 'ifany')

id_temp<-RFD4682_e$SERO_ID[RFD4682_e$VACCINATION_STATUS %in% c("Pre-Immunization")]
View(id_temp)
max(id_temp, na.rm = T) # 2020 Nov
min(id_temp, na.rm = T)

# Conclusion: The data quality of this part is good, but we cannot use those in our project
```

## The Whole APL dataset
```{r}
# The NAAT status variable
table(RFD4682_e$NAAT_STATUS, useNA = 'ifany')

# The Vaccination Status Variable
table(RFD4682_e$VACCINATION_STATUS, useNA = 'ifany')

# Cross-tab of those two
table(RFD4682_e$NAAT_STATUS, RFD4682_e$VACCINATION_STATUS, useNA = 'ifany')
```
## Categorize the time since last NAAT test
```{r}
hist(RFD4682_e$DAYS_SINCE_NAAT, breaks=20)

# Create an indicator
RFD4682_e$DAYS_NAAT_cat<-rep(NA, 214780)
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>=0 &
                        RFD4682_e$DAYS_SINCE_NAAT<=90 ]<-1  # <3 months
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>90 &
                        RFD4682_e$DAYS_SINCE_NAAT<=180 ]<-2  # 4-6 months
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>180 &
                        RFD4682_e$DAYS_SINCE_NAAT<=270 ]<-3  # 4-6 months
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>270 &
                        is.na(RFD4682_e$DAYS_SINCE_NAAT)==F ]<-4  # 4-6 months
table(RFD4682_e$DAYS_NAAT_cat, useNA = 'ifany')

# Cross-tabulation between NAAT and NAAT_DAYS categories
# Anti-N groups
table(RFD4682_e$DAYS_NAAT_cat[is.na(RFD4682_e$`N-IgG_RESULT`)==F], 
      RFD4682_e$NAAT_STATUS[is.na(RFD4682_e$`N-IgG_RESULT`)==F], useNA = 'ifany')

# Anti-RBD groups
table(RFD4682_e$DAYS_NAAT_cat[is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F &
                              RFD4682_e$VACCINATION_STATUS %in% c('Not Immunized')], 
      RFD4682_e$NAAT_STATUS[is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F &
                              RFD4682_e$VACCINATION_STATUS %in% c('Not Immunized')], useNA = 'ifany')
```

## Vaccination Status
```{r}
table(RFD4682_NAAT$VACCINATION_STATUS, useNA = 'ifany')
hist(RFD4682_NAAT$DAYS_SINCE_IMM)

# Create day gap

## NAs in NAAT Days var
table(is.na(RFD4682_NAAT$DAYS_SINCE_NAAT)) # Great, no NAs

## NAs in Vaccination Days var
table(is.na(RFD4682_NAAT$DAYS_SINCE_IMM)) # Only 911 missing, not match with Vacc Status

## Calculate the Gap
RFD4682_NAAT$DAYS_Gap<-rep(NA,6165)
RFD4682_NAAT$DAYS_Gap<-RFD4682_NAAT$DAYS_SINCE_NAAT - RFD4682_NAAT$DAYS_SINCE_IMM
table(is.na(RFD4682_NAAT$DAYS_Gap)) #911 Missing, 0 vaccinated at the same day of sample collection
hist(RFD4682_NAAT$DAYS_Gap, 
     main = 'Gap between NAAT and Vaccination',
     xlab = 'Days')

## Create the indicator
RFD4682_NAAT$Vacc_status<-rep(NA, 6165)
RFD4682_NAAT$Vacc_status[RFD4682_NAAT$VACCINATION_STATUS %in% c("Not Immunized")]<-1
RFD4682_NAAT$Vacc_status[RFD4682_NAAT$DAYS_Gap>=0 & is.na(RFD4682_NAAT$DAYS_Gap)==F]<-2
RFD4682_NAAT$Vacc_status[RFD4682_NAAT$DAYS_Gap<0 & is.na(RFD4682_NAAT$DAYS_Gap)==F]<-3
table(RFD4682_NAAT$Vacc_status, useNA = 'ifany')
```