### 2023-12-05 2_scripts/2.2 Rogen-Gladen Adj.qmd
Add RG Adj Scripts

---
title: "2.2 Rogen-Gladen Adjustment"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

## Load dataset
```{r}
# The CBS Ab subset data
load(file = "../1_data/private/CBS_Ab.RData")
load(file = "../1_data/private/RFD4682e.RData")
```

## Load necc paks
```{r}
library(dplyr)         # Data management
library(reshape2)      # for melt() function
library(ggplot)        # ggplot
library(zoo)           # year-month format
```
# ============= Part I. The Changing of Anti-N Pos APL ===============#

#### Calculate positivity directly

```{r}
# Calculate positivity for each month
tb1<-RFD4682_e[is.na(RFD4682_e$`N-IgG_RESULT`)==F,] %>%
  group_by(Col_Month) %>%
  summarise(n = n())

tb2<-RFD4682_e[RFD4682_e$`N-IgG_INTERP`=="Positive",] %>%
  group_by(Col_Month) %>%
  summarise(n.pos = n())

tb1<-merge(tb1, tb2, by = "Col_Month", all.x = T, all.y = F)
tb1$prop.n<-tb1$n.pos/tb1$n
tb1$Col_Month<-as.yearmon(tb1$Col_Month)
tb1<-tb1[order(tb1$Col_Month),]

# Add R-G adjusted prevalances
# Sen=0.844, Spec = 1.0
# Source: 
tb1$prop.n_RG<- (tb1$prop.n+1-1)/(1+0.844-1)

# remove NAs
tb1<-tb1[!is.na(tb1$Col_Month),]

tb1 %>%
  with(plot(y=.$prop.n, x = .$Col_Month, type="b",
            ylab="positivity", xlab="Month",col="blue")
       )
```

#### Apply ggplot, plot two together
```{r}
APL_melt<-melt(select(tb1, Col_Month, prop.n, prop.n_RG), id.vars = "Col_Month")

APL_melt %>%
  ggplot(aes(x=Col_Month, y=value, colour=variable)) +
  geom_point(aes(colour=variable)) +
  geom_line(aes(colour=variable))  +
  theme_bw()
```

# ============= Part II. The Changing of Anti-N Pos Overtime ===============#

#### Anti-N in CBS data (Roche)
```{r}
CBS_Ab %>%
  with(summary(roche_n))

hist(log10(CBS_Ab$roche_n))

table(CBS_Ab$interp_roche_n, useNA = 'ifany')
# 54713/(124008-1) = 44.12% much higher than APL

# Negative Cases: roche_n<1, clearly 1.0 is the COI, as it published by Roche
CBS_Ab %>%
  .[.$interp_roche_n==0, ] %>%
  with(summary(roche_n))

# Positive Cases: roche_n>=1
CBS_Ab %>%
  .[.$interp_roche_n==1, ] %>%
  with(summary(roche_n))
```

```{r}
# Check the data collection time

# Abbott only - not exist
CBS_Ab %>%
  .[!is.na(.$abbott_n) & is.na(.$roche_n), ] %>%
  with(summary(sampledate))

# Abbott & Roche
CBS_Ab %>%
  .[!is.na(.$abbott_n) & !is.na(.$roche_n), ] %>%
  with(summary(sampledate))

# Roche only
CBS_Ab %>%
  .[is.na(.$abbott_n) & !is.na(.$roche_n), ] %>%
  with(summary(sampledate))
```
## Making result table for ggplot
```{r}
# prepare melt data
tb1<-CBS_Ab%>%
  group_by(month_year) %>%
  summarise(n = n(), )

tb2<-CBS_Ab[CBS_Ab$interp_roche_n==1,] %>%
  group_by(month_year) %>%
  summarise(n.pos = n(), )

tb1<-merge(tb1, tb2, by = "month_year", all.x = T, all.y = F)
tb1$prop.n<-tb1$n.pos/tb1$n
tb1$month_year<-as.yearmon(tb1$month_year)
tb1<-tb1[order(tb1$month_year),]

# Add R-G adjusted prevalances
# Sen=99.8%, Spec = 99.5%
# Source:https://www.rochecanada.com/media/covid-19--an-additional-contribution-to-the-testing-landscape-in
tb1$prop.n_RG_h<- (tb1$prop.n+0.995-1)/(0.995+0.998-1)

# Third-party sens & spec
# Sen=90.6, spec = 100 
# Source: https://meridian.allenpress.com/aplm/article/145/1/32/443501/Comparative-Clinical-Evaluation-of-the-Roche
tb1$prop.n_RG_l<- (tb1$prop.n+1-1)/(1+0.906-1)
```

#### Apply ggplot, plot two together
```{r}
CBS_melt<-melt(select(tb1, month_year, prop.n, prop.n_RG_h, prop.n_RG_l), id.vars = "month_year")

CBS_melt %>%
  ggplot(aes(x=month_year, y=value, colour=variable)) +
  geom_point(aes(colour=variable)) +
  geom_line(aes(colour=variable))  +
  geom_hline(yintercept=0, colour='red', linetype=2) +
  theme_bw() + 
  scale_fill_discrete(name = "", labels = c("Unadjusted", "RG_adjusted: Roche", "RG_adjusted: Other study")) +
  theme(legend.title = element_blank())
```

#### Plot CBS and APL all together
```{r}
# Create a merged data
APL_melt$data<-rep("APL", length(APL_melt$Col_Month))
colnames(APL_melt)<-c("month_year", "variable", "value" , "data")

CBS_melt$data<-rep("CBS", length(CBS_melt$month_year))
# keep only RG_I (sens=0.9)
CBS_melt1<-CBS_melt %>% .[.$variable!="prop.n_RG_h",]
CBS_melt1$variable<-as.character(CBS_melt1$variable)
CBS_melt1$variable[CBS_melt1$variable!="prop.n"]<-'prop.n_RG'

APLBS<-rbind.data.frame(APL_melt, CBS_melt1)

# Do ggplot
p01<-APLBS %>%
  ggplot(aes(x=month_year, y=value, colour=variable)) +
  geom_point(aes(colour=variable, shape=data)) +
  geom_line(aes(colour=variable, linetype=data))  +
  geom_hline(yintercept=0, colour='red', linetype=2) +
  xlab("Time") + ylab("Estimated Prevalence") +
  theme_bw()

p01$labels$colour<-"Adjustment"

p01
```

## What's the reason