### 2024-02-26 5_manuscript/manuscript.qmd
Update the methods section for prob models, also add bullitins for each paragraphs to be added

---
title: "Assessing methods to adjust for assay differences across COVID-19 serological surveillance studies in Canada"
execute: 
  echo: false
format:
  docx: 
    reference-doc: word-style-reference-manuscript.docx
  html: default
bibliography: bib.bib
csl: american-medical-association-brackets.csl
editor: visual
crossref: 
  thm-title: "Table"
  thm-prefix: "Table"
  cor-title: "Table S"
  cor-prefix: "Table S"
  lem-title: "Figure S"
  lem-prefix: "Figure S"
  fig-title: "**Figure**"
---

```{=html}
<!---

NOTES:
-   To render both word and HTML, can type this into terminal

quarto render 5_manuscript/manuscript.qmd

-   There are limits in cross-referencing in Quarto. The table
   Caption doesn't work well with flextable package, and 
  There's no option for secondary (supplemental) series of 
  tables or figures. As workaround, we repurpose other
  series.I'm using
        @thm- for tables (Table X)
        @cor- for supplemental tables (Table SX)
        @lem- for supplemental figures (Figure SX)
  The space between "S" and the number for supplemental figures seems unavoidable
   for now. Can manually correct before submitting to journal.

Flextable package isn't playing nicely with quarto for crossreferencing at the 
moment but it's being actively worked on. To install latest package version, use 
  devtools::install_github("davidgohel/flextable")

--->
```
Jiacheng Chen^1,2^, Yuan Yu^1^, Sheila O'Brien^3^, Yu Nakagama^4^ ..., W. Alton Russell^1,2^

<br>

^1^School of Population and Global Health, McGill University, Montreal, Canada

^2^COVID-19 Immunity Task Force

^3^Canadian Blood Services

^4^Graduate School of Medicine, Osaka City University

<br>

**Corresponding author:**

**Key words:**

**Running title:**

{{< pagebreak >}}

```{r setup, include=FALSE}
library(ggplot2)       #plots
library(data.table)    #for using datatables instead of frames
library(scales)        #formatting plot legends and text
library(readxl)        #read.excel
library(flextable)     #generating tables
# library(ftExtra)
# library(officedown)  #formatting for word
# library(officer)
# library(stringr)
theme_set(theme_bw())
```

```{r, include=FALSE}
# Load a sample dataset
df_samp <- read_excel("../1_data/tables.xlsx", sheet = "sample")

# This table to summarize three dataset we used
df_dataset <- read_excel("../1_data/tables.xlsx", sheet = "dataset")
```

# Abstract

**Background:** A

**Methods:** A

**Results:** A

**Conclusions:** A

{{< pagebreak >}}

# Introduction

Test of citation: [@Langham2018a]. See @fig-sample, @cor-sample, @thm-sample, @thm-data, and @lem-sample.

The Covid-19 pandemic overview, global/national (in Canada) trend and how it envolved overtime in each regions of Canada. Serosurveillance programs in Canada, and its importance.

Two types of antibodies to be detected by assays: anti-spike and anti-nucleocapsid. Why we treated them differently in data analysis, one for quantitative and anothor one for qualitative test results.

-   Anti-S (humoral response to infection or vaccination) and anti-N (marker of natural infection)

-   Anti-S level elevates after infection and can be boosted by vax, while anti-N level elevates with infection (even drops) but didn't change so much with vax.

-   Anti-S level wanes slower than anti-N level after infection

Previous studies have found an association between seropositivity/ab levels and intensity of immunity (humoral + cellular). Declining ab level suggesting waning of immunity against SARS-COV-2 after infection and immunization.

Seropravalence and rolling incidence, why we need to employ RG equation or Bayesian model to estimated seropravalence, and why we also need probabilistic models to predict past infection using early 2020 and pre-Covid neg ref data.

{{< pagebreak >}}

# Methods

```{=html}
<!---

NOTES:
- The methods section of this project is partly unfinished, and the details of it needs constantly updating overtime, as the project goes. The potential obstacles were listed below:

1. APL assay data before the Omicron wave was mostly missing, only Omicron wave data is available. Also we do not know what assay they used for anti-N (probably Abbott anti-N).

2. In AB Outpatient Lab Data, they changed the assay (Diasorin S -- Abbott S) in the midway, this would affect the validity of analysis based on quantitative test results, as the results from different assay are not directly comparable. Hopefully this can be clarified by the APL data team. Our currently analysis assume the Anti-N/S assays used in omicron wave are all Abbott's, based on the paper published by Carmen in the post-vax stage: https://doi.org/10.1080/23744235.2022.2080250

3. Canpath: At least two types of anti-N assays are utilized: Mount Sinai Hosp anti-N (a in-lab developed assay) and Abbott anti-N (a commercial assay). Need to figure out which prov at what time used what assay, so we can apply sens for adjustment accordingly.

4. CBS: The data we have least problem with. CBS is still actively updating (last ver downloaded on Nov 30th) and Yuan has the access to the most updated ver of CBS. (last checked on 21-Feb-2024, CBS 2023-11-30 is still the last version)
--->
```
## Dataset

To estimate the SARS-CoV-2 seroprevalence and rolling incidence in Alberta, we used data from the Alberta Outpatient Lab (APL), Canadian Blood Services (CBS) and Canadian Partnership for Tomorrow's Health (CANPATH). Alberta Health Services (AHS) tested \>160,000 APL blood draws for SARS-CoV-2 anti-N and anti-S antibodies from June 2020 until October 2022. CBS has tested \>450,000 blood donations for both antibodies levels since March 2020. CANPATH has provided data with sample size \> XX. Besides of assay results, all three datasets include demographic variables such as donor's sex, self-reported race/ethnicity, and age. From their address, neighborhood material/social deprivation and urban/rural indicator have been derived.

The three datasets feature different data collection time frames. The CBS dataset has the most extended period, ranging from December 2020 to February 2023. APL's data collection spans from June 2020 to October 2022, but it's important to note that anti-N assay results are only available starting from December 2021. Meanwhile, the CANPATH dataset is accessible for a shorter duration, from September 2020 to May 2021.

We did not conduct analysis using a merged data, but instead, we conducted analysis separately for each data and presented the results in a graph panel (@fig-seroprevalence). @thm-data below summarized the differences among three data utilized in the study.

@thm-data CBS, CANPATH and APL

[insert @thm-data here if necessary]

## **Seroprevalence and rolling incidence**

To evaluate the severity of Covid-19 pandemic overtime in Alberta province, first we calculated the anti-N assay test positivity. Then, to bridge the gap between test positivity and population-level seroprevalence, we employed the Rogen-Gladen (RG) equation and Bayesian models.

Following seroprevalence estimation, we further estimated monthly rolling incidence proportion of SARS-CoV-2 infection. We developed a multinomial logistic regression model predicting the possibility of past infection based on the anti-N assay results, adjusting for age and gender. Then we estimated the probability of past infection in a monthly scale using population averages.

We presented both seroprevalence and rolling incidence proportion in a unified graph (@fig-seroprevalence), displaying changes over time on a monthly scale.

Rogan Gladen equation:

$$
Prevalence = \frac{TestPositivity + (Spec-1)}{Spec + (Sens-1))}
$$ Bayesian models adjusting for sensitivity:

\$\$

\$\$

Bayesian models adjusting for waning immunity:

\$\$

\$\$

## **NT50 and corresponding antibody level:**

In evaluating whether participants possess humoral immunity levels sufficient to shield them from future infections, from either vaccination or infection, we applied a threshold of protection to the quantitative Anti-S assay results. The threshold is cited from the package insert of each assays \[Roche and Abbott\], indicating sufficient antibody level to neutralize over 50% virus, with certain times of dilution (1:20). We also presented the fluctuation of the proportion sufficiently protected over time in another figure (@fig-immunity), and we conducted sensitivity analyses in which we compared alternative threshold levels. The details of the sensitivity analysis is presented in the **Fig**. Y (haven't scratched this one yet).

[Insert @fig-immunity here if necessary]

## Attack rates - TBD

The cumulative attack rates is no longer of interest, as the majority of the Canadian population has been infected at least once since the Omicron wave. As the comparison rolling incidence proportion is a measurement with more meaningful public health implications.

## Mixture  Models - TBD

In mixture models, we assume the dist of certain variables (in our case it is S/Co of Anti-N/S assay results) has two (or maybe three) distinct (though partially overlapping) distributions. Mixture models were utilized by previous studies to estimate cumulative incidence.

<https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1008728>

<https://www.sciencedirect.com/science/article/pii/S175543652200024X>

<br>

# Results

A paragraph about test sample positivity, Rogen-Gladen adjusted seroprevalence and Bayesian model adjusted seroprevalence. Describe the trend of changing overtime, difference and consistency among different adjusting methos, regions and datasets. Cite figures accordingly if necessary.

A paragraph about estimated proportion of ever infected & infected in past 3 month.

A paragraph about estimated proportion with sufficient humoral immunity to protect them from future possible SARS-CoV-2 infection. This proportion is also adjusted by Rogen-Gladen equation and Bayesian models.

<br>

# Discussion

The gap between Rogen-Gladen adjusted sero-prevalence and unadjused sero-positivity: In our case, the sensitivity is where most variance come from. As the proof of excellent assay performance, assay manufacturer published the sensitivity and specificity of their assay. However, their test performance figures derived from hospitalized patients with more severe symptoms, who were samples at the time when sero antibody level reachd its highest level. Due to the representativeness issue, it is not recommended only take the test performance figures published by the assays manufacturer into consideration.

Since sens is imperfect (0.85--1.0) and specificity is mostly 100% trustworthy (\~1.0), the R-G equation can be simplified as:

$$
Prevalence = \frac{TestPositivity}{Sens}
$$

According to the equation above, the worse the sensitivity (smaller in value), the larger the gap between the adjusted and unadjusted. We found larger gaps among the Abbott assay results, as the sensitivity we used for Abbott assays are smaller, when compared to sensitivity we used for Roche assays.

```{=html}
<!---
Notes: 
Variances among data (APL, CBS, CANPATH) is not the major focus of this study, Yuan's paper will focus on those by evaluating the representativeness of the serosurveillance data. For this paper, the major focus should be the difference between adjusting methods. The potential readers of this papers includes those who are using serosurveillance data to monitor the progress of Covid pandemic, or other pathogens of interests which possess similar features as SARS-CoV-2.

So we won't discuss the among-dataset differences on measurements in the discussion section
--->
```

Difference/Consistence among adjusting methods, regions in Canada. Is our finding consistent with previous studies?

Strength and Limitation

Public health implications:
1.  Without any adjustment, the public health inference on test positivity is very limited. It only estimates the proportion of test positive results among blood samples collected in the serosurveillance study.
2.  Adjusting for certain factors will change the public health inference, meanwhile it will also improve its generalizability. When the imperfect nature of the seroassays is taken into account, the adjusted proportion of positive test results (seroprevalence) will be able to represent the target population in our target population, from which our sample has been collected. Upon the seroprevalence, after adjusting for seroreversion using Bayesian models, we get attack rates of SARS-CoV-2, which estimates the proportion of population infected by the virus since the beginning of pandemic.
3.  There's some differences among all those different quantitative measurements, and also they have different public health inference. Future researchers should be cautious when making references to those different measurements.

<br>

{{< pagebreak >}}

# Declarations

**Funding:** A

**Conflicts:** A

**Ethics/Consent:** A

**Data and materials:** A

**Code availability:** A

**Authors' contributions: A**

{{< pagebreak >}}

# References

::: {#refs}
:::

{{< pagebreak >}}

# Tables

::: {#thm-sample}
This is a sample table

```{r}
t_samp <- as_flextable(as_grouped_data(df_samp,
                                              groups = "subhead"))
t_samp <- compose(t_samp, i = ~ !is.na(subhead), j = "col1",
              value = as_paragraph(as_chunk(subhead)))
t_samp <- fontsize(t_samp, size = 10, part = "all")
t_samp <- font(t_samp, fontname = "Times", part = "all")
t_samp <- theme_box(t_samp)
t_samp <- bg(t_samp, bg = "#EAEAEA", part = "header")
t_samp <- width(t_samp, 1, 1.7)
t_samp <- bg(t_samp, i = c(1, 4), bg = "#DDDDDD", part = "body")
t_samp <- bold(t_samp, i = c(1, 4), part = "body")
t_samp <- align(t_samp, align = "left", part = "all")

t_samp
```
:::

::: {#thm-data}
The table below displays the distribution of age groups and gender across all three datasets. It also details the provinces where the data was collected, along with the corresponding data collection periods.

```{r}
t_dataset<-as_flextable(df_dataset)
t_dataset <- fontsize(t_dataset, size = 10, part = "all")
t_dataset <- font(t_dataset, fontname = "Times", part = "all")

t_dataset <- bold(t_dataset, i = 1, part = "header")
t_dataset <- bold(t_dataset, i = c(2, 6, 9), part = "body")

#t_dataset <- theme_box(t_dataset)
#t_dataset <- bg(t_dataset, bg = "#EAEAEA", part = "header")
#t_dataset <- width(t_dataset, 1, 1.7)
#t_dataset <- bg(t_dataset, i = c(1, 4), bg = "#DDDDDD", part = "body")
#t_dataset <- bold(t_dataset, i = c(1, 4), part = "body")
#t_dataset <- align(t_dataset, align = "left", part = "all")

t_dataset
```
:::

{{< pagebreak >}}

# Figures

```{r}
#| label: fig-sample
#| fig-cap: "Figure caption here."

knitr::include_graphics("../4_output/figs/figure.png")
```

```{r}
#| label: fig-seroprevalence
#| fig-cap: "The figure panel above presents the temporal changes in seroprevalence and attack rate, grouped by province and data source. Each figure displays the unadjusted seropositivity, the seroprevalence adjusted using the Rogan-Gladen equation, the seroprevalence adjusted by Bayesian models, and the attack rates. The calculations of seropositivity, seroprevalence, and attack rate are based on the qualitative results of the anti-N assay. 
#| (Just a scratch, not the real fig)"

knitr::include_graphics("../4_output/figs/Figure.1 Seroprelance and attack rate by province and data source.png")
```

```{r}
#| label: fig-immunity
#| fig-cap: "The figure panel above depicts the temporal fluctuations in quantitative anti-S assay results, grouped by province and data source. Two cutoff indices (COIs) are utilized for different public health interpretations. The COI of 0.8, recommended by the manufacturer (Roche), is optimal for identifying the presence of antibodies in the blood sample. Meanwhile, a COI of 15 is indicative of a threshold for sufficient humoral immunity against the SARS-CoV-2 virus. We also applied Rogen-Gladen equation and Bayesian models to adjust for test sensitivity.
#| (Just one fig, will add more to build a panel)"

knitr::include_graphics("../4_output/figs/Figure.2 Immunity.png")
```

{{< pagebreak >}}

# Supplemental materials

<br>

# A. Supplement section

## Supplement Methods

Besides of serology test results and demographic data, CBS has linked the administrative data for the cohort, which provided additional information on their hospitalizations, clinic visits, diagnoses, COVID vaccinations, and long-term disability claims. Similarly, AHS and CANPATH have completed linkages and included additional vaccination history.

```{=html}
<!---
Notes: Current ver of APL has completed very limited data linkage, which only includes PCR test results/time and vaccination time. The paragraph briefly describe the data linkage for three datasets, however for now we do not need those variables from linked administrative data.
--->
```

A short paragraph about weighting: Our analysis is based on individual level. As we are calculating measurements at a monthly scale, samples sampled at different months were unweighted as they contributed independently to each month's point prevalence estimates. While in all of those three dataset, it's possible to have one individual being sampled multiple times within the same month, in this case we applied weights to each individual according to the number of samples they have in the same month.

{{< pagebreak >}}

# Supplemental tables

::: {#cor-sample}
Table caption here.

```{r}
t_samp <- as_flextable(as_grouped_data(df_samp,
                                              groups = "subhead"))
t_samp <- compose(t_samp, i = ~ !is.na(subhead), j = "col1",
              value = as_paragraph(as_chunk(subhead)))
t_samp <- fontsize(t_samp, size = 10, part = "all")
t_samp <- font(t_samp, fontname = "Times", part = "all")
t_samp <- theme_box(t_samp)
t_samp <- bg(t_samp, bg = "#EAEAEA", part = "header")
t_samp <- width(t_samp, 1, 1.7)
t_samp <- bg(t_samp, i = c(1, 4), bg = "#DDDDDD", part = "body")
t_samp <- bold(t_samp, i = c(1, 4), part = "body")
t_samp <- align(t_samp, align = "left", part = "all")

t_samp
```
:::

{{< pagebreak >}}

# Supplemental figures

::: {#lem-sample}
Figure caption here.

```{r}
knitr::include_graphics("../4_output/figs/figure.png")
```
:::