### 2024-03-05 2_scripts/2.1.1 Canpath Data.qmd
Updated codes for model w age + sex

---
title: "2.1.1 Canpath Data Check"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

The aim of this qmd file is cleaning the Canpath data. Those code is unnecessary to publish with the project.

## Running Code

## Load packages
```{r}
library(dplyr)         # Data management
library(zoo)           # year-month format
```

## Load the dataset
```{r}
# The Questionnaire data
Canpath_qx <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Qx_96014par_1125var.csv")

# The linked administrative data
Canpath_admin <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Admin_25727par.csv")

# The serology data
Canpath_ser <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Results_74503par.csv")
```