### 2024-03-07 2_scripts/2.1.1 Canpath Data.qmd
Data quality check for Canpath data

---
title: "2.1.1 Canpath Data Check"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

The aim of this qmd file is cleaning the Canpath data. Those code is unnecessary to publish with the project.

## Running Code

## Load packages

```{r}
library(dplyr)         # Data management
library(zoo)           # year-month format
```

## Load the dataset

```{r}
# The Questionnaire data
Canpath_qx <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Qx_96014par_1125var.csv")

# The linked administrative data
Canpath_admin <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Admin_25727par.csv")

# The serology data
Canpath_ser <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Results_74503par.csv")
```

##=============== Part.I Clean, merge and create a small subset===============##

#### 1.1 We need those variables from each dataset:

-   The questionnaire (qx) data:

```{r}
# age, pid, gender, location (FSA)

# age var:
summary(Canpath_qx$C_SDC_AGE) # N=3559 missing, our population is old median=64
hist(Canpath_qx$C_SDC_AGE)    # skewed normal dist, lack of young adults (18-35)

# sex/gender: use sex here
summary(Canpath_qx$C_SDC_SEX) # N=3559 missing
table(Canpath_qx$C_SDC_SEX, useNA = 'ifany')
# 0=male, 1=female. Much more f than m.

# location - 3-digit FSA
summary(Canpath_qx$C_ADM_FSA) # N=4569 missing, ~5%
table(is.na(Canpath_qx$C_ADM_FSA), useNA = 'ifany')

# province - not available, we can create one with FSA
```

-   The admin data: info about sampling

```{r}
# sample ID, 
summary(Canpath_admin$C1_SAMPLE_ID)

# cohort
summary(Canpath_admin$C1_ADM_COHORT_ID)

# date of collection
summary(Canpath_admin$C1_ADM_COLLECT_DATE)

# sample type
summary(Canpath_admin$C1_SAMPLE_TYPE)

# date of procession
summary(Canpath_admin$C1_SAMPLE_PROCESS_DATE)
```

-   The serology data: sero assay results
```{r}
# assay ID (three Sinai Hosp in-lab, 2 Abbott)
table(Canpath_ser$C1_CITF_ASSAY_ID, useNA = 'ifany')
# 10010-00: Gingras anti-N
# 10020-00: Gingras anti-S
# 10030-00: Gingras anti-S-RBD
# 20010-03: Abbott anti-N
# 20012-03: Abbott anti-S-RBD

# assay COI
table(Canpath_ser$C1_SAMPLE_ANTIGEN_CUTOFF, useNA = 'ifany')
table(Canpath_ser$C1_SAMPLE_ANTIGEN_CUTOFF, Canpath_ser$C1_CITF_ASSAY_ID)
# one single COI was applied to each assay, excellent!

# what is max value (upper dynamic range) for each assay?
table(Canpath_ser$C1_SAMPLE_ANTIGEN_MAXIMUM, useNA = 'ifany')
# not quite useful

# what is the assay results?
summary(Canpath_ser$C1_SAMPLE_RESULT, useNA = 'ifany')
hist(Canpath_ser$C1_SAMPLE_RESULT[Canpath_ser$C1_SAMPLE_RESULT<5])

# Interpretation of the results 1/0
table(Canpath_ser$C1_SAMPLE_RESULTS_DESCRIPTION, useNA = 'ifany')
```

-   All for three: pid (participants ID) and sid (sample ID)
```{r}
# sample_ID

# Sample ID
# admin data
summary(Canpath_admin$C1_SAMPLE_ID)
table(is.na(Canpath_admin$C1_SAMPLE_ID), useNA = 'ifany') # N=8349 missing, be careful
length(unique(Canpath_admin$C1_SAMPLE_ID))                # all non-missing has unique sample ID

# ser data
summary(Canpath_ser$C1_ADM_SPECIMEN_ID)
table(is.na(Canpath_ser$C1_ADM_SPECIMEN_ID), useNA = 'ifany')  # N=13632 missing, be careful
length(unique(Canpath_ser$C1_ADM_SPECIMEN_ID))                 # not all non-missing has unique sample ID

# Research ID
# qx data:
summary(Canpath_qx$ResearcherID)                        # N=96014, no missing
table(is.na(Canpath_qx$ResearcherID), useNA = 'ifany')
length(unique(Canpath_qx$ResearcherID))                 # all 96014 rid are unique

# admin data:
summary(Canpath_admin$ResearcherID)                     # N=25727, no missing
table(is.na(Canpath_admin$ResearcherID), useNA = 'ifany')
length(unique(Canpath_admin$ResearcherID))              # N=25724, duplicates

  ## find duplicates
  n_admin <- data.frame(table(Canpath_admin$ResearcherID))
  n_admin[n_admin$Freq > 1,]          # 3 duplicates, sample pp with two samples (diff sample ID)

# ser data
summary(Canpath_ser$ResearcherID)                      # N=74503, no missing
table(is.na(Canpath_ser$ResearcherID), useNA = 'ifany')
length(unique(Canpath_ser$ResearcherID))               # N=25975, duplicates
```
#### 1.2 Merge three data
```{r}
# select the variables we need from each subset:
# qx subset:
qx_s<-select(Canpath_qx, ResearcherID, C_ADM_STUDY_DATASET, C_SDC_AGE, C_SDC_SEX,
             C_ADM_FSA)
# apply shorter var names:
colnames(qx_s)<-c("rid", "study_id", "age", "sex", "fsa")

# admin subset
admin_s<-select(Canpath_admin, ResearcherID, C1_SAMPLE_ID,
                C1_ADM_COHORT_ID, C1_ADM_COLLECT_DATE, C1_SAMPLE_TYPE, 
                C1_SAMPLE_PROCESS_DATE)
colnames(admin_s)<-c("rid", "sid", "cohort_id", "date_col", "stype", "date_pro")

# ser data, use anti-N only
ser_s<-select(Canpath_ser[Canpath_ser$C1_CITF_ASSAY_ID %in% c("10010-00", "20010-03"),],
              ResearcherID, C1_ADM_SPECIMEN_ID, C1_CITF_ASSAY_ID, C1_SAMPLE_DATE_ASSAY_RUN,
              C1_SAMPLE_RESULT, C1_SAMPLE_RESULTS_DESCRIPTION)
colnames(ser_s)<-c("rid", "sid", "assay_id", "date_asy", "result", "pos")

# reshape the subset
# ser_sw<-reshape(ser_s, idvar = "sid", timevar = "assay_id", direction = "wide")
# reshape is unnecessary
```
```{r}
# merge those small subsets
m1<-merge(admin_s, ser_s[!is.na(ser_s$rid) & !is.na(ser_s$sid),],
          by = c("rid", "sid"), all.x = F, all.y = T)
m2<-merge(m1, qx_s, by='rid', all.x = T, all.y = F)

ser_s %>%
  .[!is.na(.$rid) & is.na(.$sid),]

# is cohort_id same as study_id?
table(m2$cohort_id, m2$study_id, useNA = 'ifany')
# looks messy, use FSA to determine prov of residency then

# unique participants (rid)
summary(m2$rid)
table(is.na(m2$rid), useNA = 'ifany')          # N=21536, no missing

# unique samples (sid)
summary(m2$sid)
table(is.na(m2$sid), useNA = 'ifany')          # N=21536, no missing
length(unique(m2$sid))                         # N=21349, some pp has multiple samples
```
#### Apply FSA-province conversion
```{r}
# this chunk is provided by Y.Yu
province_fun <- function(var) {
  fsa_f = as.character(substr(var,1,1))
  prov =  ifelse(fsa_f == "A", "NL",
                 ifelse(fsa_f == "B", "NS",
                        ifelse(fsa_f == "C", "PE",
                               ifelse(fsa_f == "E", "NB",
                                      ifelse(fsa_f == "G" | fsa_f == "H" | fsa_f == "J", "QC",
                                             ifelse(fsa_f == "K" | fsa_f == "L" | fsa_f == "M" | fsa_f == "N" | fsa_f == "P", "ON",
                                                    ifelse(fsa_f == "R", "MB",
                                                           ifelse(fsa_f == "S", "SK",
                                                                  ifelse(fsa_f == "T", "AB",
                                                                         ifelse(fsa_f == "V", "BC",
                                                                                ifelse(fsa_f == "X", "NU/NT",
                                                                                       ifelse(fsa_f == "Y", "YT", NA
                                                                                       ))))))))))))

  return(prov)}

# apply the function
m2$prov<-province_fun(m2$fsa)
table(m2$prov, m2$study_id, useNA = 'ifany')

# need to clean NA's
table(m2$cohort_id, m2$study_id, useNA = 'ifany')
table(m2$prov, m2$cohort_id, useNA = 'ifany')

## 1) group by prov (FSA converted)
  m2$region[m2$prov %in% c("NB", "NL", "NS", "PE")]<-"Atlantic"
  m2$region[m2$prov %in% c("AB")]<-"Alberta"
  m2$region[m2$prov %in% c("BC")]<-"BC"
  m2$region[m2$prov %in% c("ON")]<-"Ontario"
  m2$region[m2$prov %in% c("QC")]<- "QC"  # recode it as NA later
  m2$region[m2$prov %in% c("SK")]<- "SK & MB"
table(m2$prov, m2$region, useNA = 'ifany')

## 2) clean NA, use study_id
table(m2[is.na(m2$region),]$study_id, 
      m2[is.na(m2$region),]$cohort_id, useNA = 'ifany')
  m2$region[is.na(m2$region) &
            m2$cohort_id == 1] <- "Atlantic"
  m2$region[is.na(m2$region) &
            m2$cohort_id == 3] <- "BC"
  m2$region[is.na(m2$region) &
            m2$cohort_id == 5] <- "SK & MB"
  m2$region[is.na(m2$region) &
            m2$cohort_id == 6] <- "Ontario"
  m2$region[is.na(m2$region) & is.na(m2$cohort_id) &
            m2$study_id == "ATP"] <- "Alberta"
  m2$region[is.na(m2$region) & is.na(m2$cohort_id) &
            m2$study_id == "OHS"] <- "Ontario"
table(m2$region, useNA = 'ifany')

## recode QC as missing
  m2$region[m2$region=="QC"]<-NA # N=4 coded as missing (NA)
```
## see what is the assay used in each region
```{r}
m2 %>%
  select(region, assay_id) %>%
  group_by(region, assay_id) %>%
  summarise(Freq=n())
```
## save the data
```{r}
Canpaths<-m2
rm(m1, m2, n_admin, ser_sw)
save(Canpaths, file = "../1_data/private/CANPATH/Canpaths.RData")
```