### 2024-03-18 2_scripts/2.1.1 Canpath Data.qmd
Update fig code for Canpath

---
title: "2.1.1 Canpath Data Cleaning"
format: html
editor: visual
---

## Quarto

The aim of this qmd file is cleaning the Canpath data.

In Canpath data we've removed three individuals who provided multiple samples, and several samples tested for multiple times. Those samples tested for multiple times did not result in inconsistent interpretations on assay results. See codes below for more details on data cleaning.

In cleaned ver of Canpath data all rid (researcherID, participant ID) and sid (sample ID) are unique and 1-to-1, though some sid is missing.

## Running Code

## Load packages

```{r}
library(dplyr)         # Data management
# library(zoo)           # year-month format
```

## Load the dataset

```{r}
# The Questionnaire data
Canpath_qx <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Qx_96014par_1125var.csv")

# The linked administrative data
Canpath_admin <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Admin_25727par.csv")

# The serology data
Canpath_ser <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Results_74503par.csv")

# The new MTP Qx subset
Canpath_qx_MTP<-read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Qx_1114par_1127varMTP.csv")
```

##=============== Part.I Clean, merge and create a small subset===============##

#### 1.1 We need those variables from each dataset:

-   The questionnaire (qx) data:

```{r}
# age, pid, gender, location (FSA)

# age var:
summary(Canpath_qx$C_SDC_AGE) # N=3559 missing, our population is old median=64
hist(Canpath_qx$C_SDC_AGE)    # skewed normal dist, lack of young adults (18-35)

# sex/gender: use sex here
summary(Canpath_qx$C_SDC_SEX) # N=3559 missing
table(Canpath_qx$C_SDC_SEX, useNA = 'ifany')
# 0=male, 1=female. Much more f than m.

# location - 3-digit FSA
summary(Canpath_qx$C_ADM_FSA) # N=4569 missing, ~5%
table(is.na(Canpath_qx$C_ADM_FSA), useNA = 'ifany')

# province - not available, we can create one with FSA
```

-   The admin data: info about sampling

```{r}
# sample ID, 
summary(Canpath_admin$C1_SAMPLE_ID)

# cohort
summary(Canpath_admin$C1_ADM_COHORT_ID)

# date of collection
summary(Canpath_admin$C1_ADM_COLLECT_DATE)

# sample type
summary(Canpath_admin$C1_SAMPLE_TYPE)

# date of procession
summary(Canpath_admin$C1_SAMPLE_PROCESS_DATE)
```

-   The serology data: sero assay results
```{r}
# assay ID (three Sinai Hosp in-lab, 2 Abbott)
table(Canpath_ser$C1_CITF_ASSAY_ID, useNA = 'ifany')
# 10010-00: Gingras anti-N
# 10020-00: Gingras anti-S
# 10030-00: Gingras anti-S-RBD
# 20010-03: Abbott anti-N
# 20012-03: Abbott anti-S-RBD

# assay COI
table(Canpath_ser$C1_SAMPLE_ANTIGEN_CUTOFF, useNA = 'ifany')
table(Canpath_ser$C1_SAMPLE_ANTIGEN_CUTOFF, Canpath_ser$C1_CITF_ASSAY_ID)
# one single COI was applied to each assay, excellent!

# what is max value (upper dynamic range) for each assay?
table(Canpath_ser$C1_SAMPLE_ANTIGEN_MAXIMUM, useNA = 'ifany')
# not quite useful

# what is the assay results?
summary(Canpath_ser$C1_SAMPLE_RESULT, useNA = 'ifany')
hist(Canpath_ser$C1_SAMPLE_RESULT[Canpath_ser$C1_SAMPLE_RESULT<5])

# Interpretation of the results 1/0
table(Canpath_ser$C1_SAMPLE_RESULTS_DESCRIPTION, useNA = 'ifany')
```

-   All for three: pid (participants ID) and sid (sample ID)
```{r}
# sample_ID: only available in admin and ser data

# Sample ID
# admin data
summary(Canpath_admin$C1_SAMPLE_ID)
table(is.na(Canpath_admin$C1_SAMPLE_ID), useNA = 'ifany') # N=8349 missing, be careful
length(unique(Canpath_admin$C1_SAMPLE_ID))                # all non-missing has unique sample ID

# ser data
summary(Canpath_ser$C1_ADM_SPECIMEN_ID)
table(is.na(Canpath_ser$C1_ADM_SPECIMEN_ID), useNA = 'ifany')  # N=13632 missing, be careful
length(unique(Canpath_ser$C1_ADM_SPECIMEN_ID))                 # not all non-missing has unique sample ID

# Research ID: invididual ID
# qx data:
summary(Canpath_qx$ResearcherID)                        # N=96014, no missing
table(is.na(Canpath_qx$ResearcherID), useNA = 'ifany')
length(unique(Canpath_qx$ResearcherID))                 # all 96014 rid are unique

# admin data:
summary(Canpath_admin$ResearcherID)                     # N=25727, no missing
table(is.na(Canpath_admin$ResearcherID), useNA = 'ifany')
length(unique(Canpath_admin$ResearcherID))              # N=25724, duplicates

  ## find duplicates
  n_admin <- data.frame(table(Canpath_admin$ResearcherID))
  n_admin[n_admin$Freq > 1,]          # 3 duplicates, one pp with two samples (diff sample ID), collected on diff day. Drop the first one

# ser data
summary(Canpath_ser$ResearcherID)                      # N=74503, no missing
table(is.na(Canpath_ser$ResearcherID), useNA = 'ifany')
length(unique(Canpath_ser$ResearcherID))               # N=25975, duplicates due to same sample tested for multiple assays
```

#### 1.2 Merge three data
```{r}
# select the variables we need from each subset:
# qx subset:
qx_s<-select(Canpath_qx, ResearcherID, C_ADM_STUDY_DATASET, C_SDC_AGE, C_SDC_SEX,
             C_ADM_FSA)

# don't forget MTP subset
qx_s_mtp<-select(Canpath_qx_MTP, ResearcherID, C_ADM_STUDY_ID, 
             C_SDC_AGE, C_SDC_SEX, C_ADM_FSA,
             C1_SDC_AGE, C1_SDC_SEX, C1_ADM_FSA)
# apply shorter var names:
colnames(qx_s)<-c("rid", "study_id", "age", "sex", "fsa")
colnames(qx_s_mtp)<-c("rid", "study_id", "age", "sex", "fsa", 
                      "age1", "sex1", "fsa1")

# identify those missed from original data:
qx_s_mtp[qx_s_mtp$rid %in% qx_s$rid,] # not missed, N=699
qx_mtpmiss<-qx_s_mtp[!(qx_s_mtp$rid %in% qx_s$rid),] # missed, N=445

# rbind those two into one:
qx_s<-rbind.data.frame(qx_s, 
                       select(qx_mtpmiss, -c("age1", "sex1", "fsa1")))

# clean study_id
qx_s$study_id[qx_s$study_id=='5']<-"MTP"
table(qx_s$study_id, useNA = 'ifany')

# admin subset
admin_s<-select(Canpath_admin, ResearcherID, C1_SAMPLE_ID,
                C1_ADM_COHORT_ID, C1_ADM_COLLECT_DATE, C1_SAMPLE_TYPE, 
                C1_SAMPLE_PROCESS_DATE)
colnames(admin_s)<-c("rid", "sid_admin", "cohort_id", "date_col", "stype", "date_pro")

# ser data, use anti-N only
ser_s<-select(Canpath_ser[Canpath_ser$C1_CITF_ASSAY_ID %in% c("10010-00", "20010-03"),],
              ResearcherID, C1_ADM_SPECIMEN_ID, C1_CITF_ASSAY_ID, C1_SAMPLE_DATE_ASSAY_RUN,
              C1_SAMPLE_RESULT, C1_SAMPLE_RESULTS_DESCRIPTION)
colnames(ser_s)<-c("rid", "sid_ser", "assay_id", "date_asy", "result", "pos")

table(ser_s$rid %in% qx_mtpmiss$rid, useNA = 'ifany')
table(admin_s$rid %in% qx_mtpmiss$rid, useNA = 'ifany')
```

#### Data quality check of those small dataset
```{r}
# admin_s: missing
table(is.na(admin_s$rid), useNA = 'ifany')     # rid no missing, not all unique (3 duplicates)
# remove duplicates: drop the earliest sample
admin_s<-admin_s[!admin_s$sid %in% c("DB0006218", "DB0009025", "DB0009310"),]

table(is.na(admin_s$sid), useNA = 'ifany')     # sid 8349 missing
length(unique(admin_s$sid[!is.na(admin_s$sid)])) # non-missing sids all unique
# remove duplicates: drop the earliest sample
ser_s<-ser_s[!ser_s$sid %in% c("DB0009310"),]

# ser_s (only 2 anti-N assays): missing
table(is.na(ser_s$rid), useNA = 'ifany')      # rid no missing
length(unique(ser_s$rid))                     # duplicates exists

  n_ser <- data.frame(table(ser_s$rid))
  n_ser[n_ser$Freq > 1,]
# some sample has been tested multiple times by the same assay
# none of those testing has been result in inconsistency of anti-N positivity

# in ser_s, see where rid is duplicated: all used Mount Sinai Hosp Assay (anti-N)
ser_rid_dup<-ser_s[ser_s$rid %in% n_ser[n_ser$Freq > 1,]$Var1, ]
table(ser_s[ser_s$rid %in% n_ser[n_ser$Freq > 1,]$Var1, ]$assay_id, useNA='ifany')

# need to deduplicate ser_s, only keep one of duplicated records
ser_rid_dup$id<-c(1:402)
ser_rid_dup %>%
  distinct(rid, .keep_all=TRUE) # N=162 （402-162 = 240 de-duplicated)

ser_s<-ser_s %>%
  distinct(rid, .keep_all=TRUE) # N=25839, number matches
```

```{r}
# merge those small subsets
m1<-merge(admin_s, ser_s,
          by = c("rid"), all.x = F, all.y = T)
m2<-merge(m1, qx_s, by='rid', all.x = T, all.y = F)

# sid match
m2$sid_match[!is.na(m2$sid_admin) & !is.na(m2$sid_ser) &
             m2$sid_admin == m2$sid_ser]<-"match"
m2$sid_match[!is.na(m2$sid_admin) & !is.na(m2$sid_ser) &
             m2$sid_admin != m2$sid_ser]<-"mismatch"
m2$sid_match[!is.na(m2$sid_admin) & is.na(m2$sid_ser)]<-"ser miss"
m2$sid_match[is.na(m2$sid_admin) & !is.na(m2$sid_ser)]<-"admin miss"
m2$sid_match[is.na(m2$sid_admin) & is.na(m2$sid_ser)]<-"all miss"
table(m2$sid_match, useNA = 'ifany') # no mismtach, good

# drop unnecessary sids
m2$sid<-m2$sid_ser
m2<-select(m2, -c("sid_admin", "sid_ser", "sid_match"))

# is cohort_id same as study_id?
table(m2$cohort_id, m2$study_id, useNA = 'ifany')
# looks messy, use FSA to determine prov of residency then

# unique participants (rid)
summary(m2$rid)
table(is.na(m2$rid), useNA = 'ifany')          # N=25839, no missing
length(unique(m2$rid))                         # all unique

# unique samples (sid)
summary(m2$sid)
table(is.na(m2$sid), useNA = 'ifany')          # N=21346, 4493 missing
length(unique(m2$sid[!is.na(m2$sid)]))         # N=21346, no missing all unique
```
#### Apply FSA-province conversion
```{r}
# this chunk is provided by Y.Yu
province_fun <- function(var) {
  fsa_f = as.character(substr(var,1,1))
  prov =  ifelse(fsa_f == "A", "NL",
                 ifelse(fsa_f == "B", "NS",
                        ifelse(fsa_f == "C", "PE",
                               ifelse(fsa_f == "E", "NB",
                                      ifelse(fsa_f == "G" | fsa_f == "H" | fsa_f == "J", "QC",
                                             ifelse(fsa_f == "K" | fsa_f == "L" | fsa_f == "M" | fsa_f == "N" | fsa_f == "P", "ON",
                                                    ifelse(fsa_f == "R", "MB",
                                                           ifelse(fsa_f == "S", "SK",
                                                                  ifelse(fsa_f == "T", "AB",
                                                                         ifelse(fsa_f == "V", "BC",
                                                                                ifelse(fsa_f == "X", "NU/NT",
                                                                                       ifelse(fsa_f == "Y", "YT", NA
                                                                                       ))))))))))))

  return(prov)}

# apply the function
m2$prov<-province_fun(m2$fsa)
table(m2$prov, m2$study_id, useNA = 'ifany')

# need to clean NA's
table(m2$cohort_id, m2$study_id, useNA = 'ifany')
table(m2$prov, m2$cohort_id, useNA = 'ifany')

## 1) group by prov (FSA converted)
  m2$region[m2$prov %in% c("NB", "NL", "NS", "PE")]<-"Atlantic"
  m2$region[m2$prov %in% c("AB")]<-"Alberta"
  m2$region[m2$prov %in% c("BC")]<-"BC"
  m2$region[m2$prov %in% c("ON")]<-"Ontario"
  m2$region[m2$prov %in% c("QC")]<- "QC"  # recode it as NA later
  m2$region[m2$prov %in% c("SK", "MB")]<- "SK & MB"
table(m2$prov, m2$region, useNA = 'ifany')

## 2) clean NA, use study_id
table(m2[is.na(m2$region),]$study_id, 
      m2[is.na(m2$region),]$cohort_id, useNA = 'ifany')
  m2$region[is.na(m2$region) &
            m2$cohort_id == 1] <- "Atlantic"
  m2$region[is.na(m2$region) &
            m2$cohort_id == 3] <- "BC"
  m2$region[is.na(m2$region) &
            m2$cohort_id == 5] <- "SK & MB"
  m2$region[is.na(m2$region) &
            m2$cohort_id == 6] <- "Ontario"
  m2$region[is.na(m2$region) & is.na(m2$cohort_id) &
            m2$study_id == "ATP"] <- "Alberta"
  m2$region[is.na(m2$region) & is.na(m2$cohort_id) &
            m2$study_id == "OHS"] <- "Ontario"
table(m2$region, useNA = 'ifany')

## recode QC as missing
  m2$region[m2$region=="QC"]<-NA # N=4 coded as missing (NA)
```

#### see what is the assay used in each region
```{r}
m2 %>%
  select(region, assay_id) %>%
  group_by(region, assay_id) %>%
  summarise(Freq=n())

# clean the date var (codes provided by M.Knight)

## The collection date var
m2$date_col1<-gsub("/","-", m2$date_col) # replace "/" with "-"
m2$date_col1<-as.Date(lubridate::parse_date_time(m2$date_col1,
                                                 c("%y-%m-%d","%m-%d-%y")),tz = "UTC")

## The assay running date var
m2$date_asy[m2$date_asy=="21-07-16"]<-"2021-07-16"
m2$date_asy1<-as.Date(m2$date_asy, format = "%Y-%m-%d")

## cleaning temp var
m2$date_asy<-m2$date_asy1
m2$date_col<-m2$date_col1
m2<-select(m2, -c("date_col1", "date_asy1"))

## Beware some discrepancies: sample collection time is later than assay date
## N=4
m2[m2$date_asy < m2$date_col &
             !is.na(m2$date_asy) & !is.na(m2$date_col),]

# when those assay was collected?
m2 %>%
  group_by(region) %>%
  summarise(Freq=n(),
            tcol_min = min(date_col, na.rm=T),
            tcol_max = max(date_col, na.rm=T))
```
#### 1.3 Add additional variables
```{r}
# add race variable: the coding on race is a bit messy in Canpath
# 14 variables were used to indicate diff racial minorities
# The cleaning code is provided by M.Knight
m2<-merge(m2, select(Canpath_qx, "ResearcherID",
          "C1_SDC_EB_ARAB","C1_SDC_EB_BLACK","C1_SDC_EB_CHINESE",
         "C1_SDC_EB_FILIPINO","C1_SDC_EB_JAPANESE",
         "C1_SDC_EB_KOREAN","C1_SDC_EB_LATIN","C1_SDC_EB_S_ASIAN",
         "C1_SDC_EB_SE_ASIAN","C1_SDC_EB_W_ASIAN","C1_SDC_EB_WHITE",
         "C1_SDC_EB_OTHER", "C1_SDC_EB_OTHER_OTSP","C1_SDC_EB_CA"),
         by.x = 'rid', by.y ="ResearcherID", 
         all.x = T, all.y = F

)

# Reorder the variables to fit Matthew's code:
m3<-cbind.data.frame(m2[,1:8], m2[,17:30], m2[,9:16])
m2<-m3
rm(m3)

# If possible, add a short para in the discussion section explaining why we can not categorize the indigenous people. Consult Yuan's email for more detail.
```

```{r}
# define whites and minorities:

#Racialized minorities - no indigenous responses included (applied later in for loop)
rm_indn<-c("Afro Latino Caribbean decent","East Indian descent, born in Kenya","Indo-Caribbean","white/asian",
           "SE Asian (Taiwanese)","i am South African, I carry in my veins European, African, Asian and indigenous Southern Afircan",
           "West Indian from Guyana","Guyanese - West Indian","South American","caucasian - european, indian, iraqi",
           "Ugandian of South Asian heritage","Trinidadian","White, Chinese, West Asian", "Japanese/Irish","Chinese, Indian, English",
           "East Indian born in Guyana","West-Indian","South African of mixed races","half Japanese half White",
           "Armenian","South Asian from East Africa","Indo Caribbean","Japanese and European","Black/white",
           "Mixed race Chinese-Guyanese-White","East Asian","west indian",
           "White, Indigeneous, Latin-American","Indian born in the Caribbean.",
           "White and Khoisan","European Hispanic","Armenian mother, Irish father, but born in England",
           "Persian","Guyanese,south asian descent","Mixed heritage (white and black)","Caribbean, Mixed Race",
           "Assyrian","Caucasian African","WEST INDIAN/CHINESE/","Latino and European","Of Indian decent - Born in Caribbean",
           "White/Southeast Asian","mixed european/chinese","Indo-caribbean","Mixed Race (Black, White, Hispanic,Soth Asian)",
           "South Asian/ Mixed South african","Canadian, German & Caribbean","Black Canadian",
           "East Indian","Mixed European & Southeast Asian","Je suis Afro Latino-AmÃ\u0083Â\u0083Ã\u0082Â©ricain","latino-amÃ\u0083Â\u0083Ã\u0082Â©ricain occidental",
           "Nord africain","Berbere","Berbere (Afrique du nord)","Berbere Nord Africain","amazigh","maghrebine","Nord-Africain",
           "Juif marocain","Nord africaine","90% white 10%Jewish-Arab","Canadian and Afrikaans",
           "mixed with Jamaican", "guianese of east indian descent",
           "caribbean chinese", "chinese and french background and other",
           "arabe-africaine", "berbère","persian",
           "indo-european","colombienne","japanese & swedish",
           "east indian","middle east",
           "armenian",
           "north african","south american, african, chinese","middle eastern",
           "middle east, and central europe","cherokee",
           "irish and guyanese","african","zimbawayan",
           "dutch indonesian",
           "trinidadian","south america - biracial",
           "punjabi","west indian of east indian descent",
           "mixed race - mother part black/father white caucasian","jamaican",
           "canadian chinese","cambodgienne","africain","asian","mauricien",
           "taiwanese","middle east/ persion",
           "1/4 middle eastern (syria which is now lebanon) from paternal grandfather",
           "mixed afghan and english","tunisien",
           "jamaican.","west  indian","zoroastrian (ancient persian ancestry)",
           "white/west asian","kabyle (berbère) d'afrique du nord","white mexican jewish")

#Racialized minorities - indigenous responses included
rm_indy<-c("Metis/European descent","White and aboriginal","Metis","Indigenous / European","Metis French Canadian",
           "Metis Canadian","Blanche ascendance micmak","Metis ancestry","white a bit of indigenous","métis",
           "french metis", "30% indigenous metis/70% white","metis",
           "culturally white, 3/16 aboriginal","mik'maw","inuit / european", "autochtone",
           rm_indn)

#White or Caucasian
w<-c("white newfoundlander","White British",
     "White passing","Caucasian","White - British Decent",
     "White, British descent","Caucasian white", "White - Canadian Irish descent",
     "white french canadian", "Caucasian female","White - Canadian","caucasian","White Canadian Scottish parents","Canadian Caucasian","White British Descent",
     "Canadian - White - 4 generations",
     "white, British descent","CAUCASIAN",
     "Caucasian - European descent","Canadian white Caucasian","white Canadian" ,
     "Canadian White","White Franco Ontarian","Caucasian Canadian",
     "white  Canadian       North American", "White Canadian born","White Canadian, Parents of Ukrainian and French background",
     "White Anglo-Saxon","White/Cdn born/European grandparents",
     "Canadian, white","White caucasian","white not Europe.",
     "Hebrew (Jewish) Canadian white","White (Canadian)","White","White from Canada",
     "White  English decent=","White bilingual Canadian - Family from Quebec","white canadian","White  Canadian",
     "White Canadian, no known ethnicities","White (Causcasian)","white","blanc Canadienne","Canadienne blanche",
     "BLANCHE","Blanc.", "Blanc (Canada)","Canadienne blanc","Caucasien","Blanc - Canadien",
     "Blanc canadienne","Blanc du canada","Blanc canadien","Blanc canadian",
     "Blanche (Canadienne)","une femme de race blanche","blanc","Blanche","Blanc canada","Blanc","Blanche, Canadienne",
     "blanc quebecois","caucasien",
     "White & Canadian","White Canadian","blanc du canada","White- true blue CANADIAN","White Scottish Descent",
     "blanche canadienne","Blanc Canadien","canadien francais blanc",
     "white, scottish and other",
     "ccaucasian","caucasienne",
     "white european","white, mainly scottish/english/irish",
     "white, of european decent with mi'kmaq ancestors several generations removed.",
     "canadian caucasian", "white, norwegian and scottish",
     "canadien francais, caucasien","european caucasian",
     "white celtic")
```

```{r}
# a for loop for creating race var:
m2$race<-NULL

for(i in 1:nrow(m2)){
  m2$race[i]<-case_when(
    #select pnts and nothing else
    !is.na(m2[i,22]) & 
      sum(is.na(m2[i,9:22])) == 13 ~ "pnts",

    #select only 1 white ethnicity
    !is.na(m2[i,19]) & 
      sum(is.na(m2[i,9:22])) == 13 ~ "White",

    #select only 1 racialized minority ethnicity
    sum(!is.na(m2[i,9:18])) == 1 &
      sum(is.na(m2[i,9:22])) == 13 ~ "Racialized minority",

    #select 1 white ethnicity and at least 1 racialized minority ethnicity
    sum(!is.na(m2[i,9:18])) > 0 & 
      !is.na(m2[i,19]) &
      sum(is.na(m2[i,22])) ~ "Racialized minority",

    #select 2+ racialized minority ethnicities
    sum(!is.na(m2[i, 9:18])) > 1 &
      sum(is.na(m2[i, c(19, 22)])) == 2 ~ "Racialized minority",

    #select at least 1 racialized minority ethnicity and at least 1 "other" ethnicity
    sum(!is.na(m2[i, 9:18])) > 0 &
      sum(is.na(m2[i, c(19, 22)])) == 2 &
      sum(!is.na(m2[i, 20,21])) > 0 ~ "Racialized minority",

    #select white, but also identify as a racialized minority using text box
    sum(is.na(m2[i, c(9:18, 22)])) == 11 &
      !is.na(m2[i,19]) &
      sum(!is.na(m2[i, c(20:21)])) >= 1 &
      m2[i, c(21)] %in% (rm_indn) ~ "Racialized minority",

    #select white, but also identify as white using text box
    sum(is.na(m2[i, c(9:18, 22)])) == 11 &
      !is.na(m2[i,19]) &
      sum(!is.na(m2[i, c(20:21)])) >= 1 &
      m2[i, c(21)] %in% (w) ~ "White",

    #select other and identify as a racialized minority using text box
    sum(is.na(m2[i, c(9:19, 22)])) == 12 &
      m2[i, 21] %in% (rm_indn) ~ "Racialized minority",

    #select other and identify as white using text box
    sum(is.na(m2[i, c(9:19, 22)])) == 12 &
      m2[i, 21] %in% (w) ~ "White",

    TRUE ~ NA)
}

table(m2$race, useNA = 'ifany')

# pnt = prefer not to answer
table(m2[,22], m2$race, useNA = 'ifany')

# remove loose race vars
m2<-select(m2, -c(9:22))
```

#### 1.4 save the data
```{r}
Canpaths<-m2
rm(m1, m2, n_admin, ser_sw, n_ser, qx_mtpmiss, ser_rid_dup)
rm(admin_s, qx_s, qx_s_mtp, ser_s)
save(Canpaths, file = "../1_data/private/CANPATH/Canpaths.RData")

# Save to the shared folder 
save(Canpaths, file = "~/../../jmy/CANPATH/Canpaths.RData")
# Save to the shared folder as .csv format
write.csv(Canpaths, file = "~/../../jmy/CANPATH/Canpaths.csv", row.names=FALSE)
```