### 2024-06-11 2_scripts/2.1 APL Data Quality Check_new.qmd
Finished APL ver2 data cleaning (Jan 2024)

---
title: "2.1 APL Data Quality Check with additonal cleaning"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

## Load the APL data with encrypted ID variables

```{r}
# APL first version: provided by Carmen in 2023
load(file = "../1_data/private/RFD4682e.RData")

# Second version of APL: provided by Carmen in Jan 2024
# We load all columns as char, to avoid potential data corruption issues
library(readxl)
RFD4682_v2 <- read_excel("../1_data/private/RFD4682_Data_for release Jan 2024.xlsx", 
    sheet = "RFD4682_Full_Data", col_types = c("text", 
        "text", "text", "text", "text", 
        "date", "text", "text", "date", "date", 
        "text", "text", "numeric", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text"))
View(RFD4682_v2)

RFD4682_v2_vax <- read_excel("../1_data/private/RFD4682_Data_for release Jan 2024.xlsx", 
    sheet = "Vax Data", col_types = c("text", "numeric", "date", "text"))

RFD4682_v2_NAAT <- read_excel("../1_data/private/RFD4682_Data_for release Jan 2024.xlsx", 
    sheet = "NAAT Results", col_types = c("text", "date", "text"))

# There's some minor updates they made to the v2 (Jan 2024), so it's recommended to rerun the data cleaning code on v2, instead of adding new variables to v1 (June 2023)
```

## Load paks

```{r}
library(dplyr)         # Data management
# library(lubridate)     # Manage variables in Date formats
library(zoo)           # for yearmonth Date format
# library(reshape2)      # for melt() function
```

```{r}
## Create subset for those NAAT Positive Cases
table(RFD4682_e$NAAT_STATUS, useNA = 'ifany')
RFD4682_NAAT<-RFD4682_e[RFD4682_e$NAAT_STATUS %in% c('Positive'),]

# Save the subset

# Load Data file

```

## 1. APL v1 (June 2023) - Data Cleaning

#### 1.1 Some Desc Analysis

##### The distribution of time from NAAT to Sero-assay

```{r}
hist(RFD4682_e$DAYS_SINCE_NAAT[RFD4682_e$NAAT_STATUS=="Positive"], breaks = 18,
     main="Time form NAAT to Sero-assay (Days)",
     xlab = "Days")

## The time variable is good to go
```

##### Records are not patients, how many individual has been included?

```{r}
## ULI_ID: Unique Lifetime Identifier (ULI), numerical values of Alberta PHN.
length(unique(RFD4682_e$ULI_IDe))     # Unique = 180821
table(is.na(RFD4682_e$ULI_IDe))       # NA = 11298

## PERSON_ID: assigned from the originating LIS
length(unique(RFD4682_e$PERSON_IDe))     # Unique = 189864
table(is.na(RFD4682_e$PERSON_IDe))       # NA = 2973

## SERO_ID: Unique Specimen ID + Test month, unique APL dataset
length(unique(RFD4682_e$SERO_ID)) # N=214,772
table(is.na(RFD4682_e$SERO_ID))   # No missing

## Checking duplicated records by SERO_ID
tab_SERO_ID<-table(RFD4682_e$SERO_ID)
View(tab_SERO_ID)

str(tab_SERO_ID)
tab_SERO_ID.dat<-as.data.frame(tab_SERO_ID)
tab_SERO_ID.dat[tab_SERO_ID.dat$Freq==2,]
```

##### Why unique PERSON_ID and ULI_ID are not matching with each other?

```{r}
## Check the discrepancies between those two

## Freq tab for ULI_ID
tab_ULI_IDe<-table(RFD4682_e$ULI_IDe)
tab_ULI_IDe.dat<-as.data.frame(tab_ULI_IDe)
colnames(tab_ULI_IDe.dat)<-c('ULI_IDe','Freq_ULI')

# tab_ULI_ID.dat.2<-tab_ULI_ID.dat[tab_ULI_ID.dat$Freq>1,]

## Freq tab for RERSON_ID
tab_PERSON_IDe<-table(RFD4682_e$PERSON_IDe)
tab_PERSON_IDe.dat<-as.data.frame(tab_PERSON_IDe)
colnames(tab_PERSON_IDe.dat)<-c('PERSON_IDe','Freq_PER')

## Add freq var to the subset
RFD4682_e$order_ID<-1:214780
RFD4682_3<-select(RFD4682_e, order_ID, SERO_ID, ULI_IDe, PERSON_IDe)
RFD4682_3<-merge(RFD4682_3, tab_ULI_IDe.dat, by='ULI_IDe',
      all.x = T, all.y = F)
RFD4682_3<-merge(RFD4682_3, tab_PERSON_IDe.dat, by='PERSON_IDe',
      all.x = T, all.y = F)

## Find Discrepancies by compare the frequencies

## NA is messy with logic, replace it with 0
RFD4682_3$Freq_ULI[is.na(RFD4682_3$Freq_ULI)==T]<-0
RFD4682_3$Freq_PER[is.na(RFD4682_3$Freq_PER)==T]<-0
tab.idfreq<-table(RFD4682_3$Freq_ULI, RFD4682_3$Freq_PER, useNA = 'ifany')
tab.idfreq
dat.idfreq<-as.data.frame(tab.idfreq)

# Scenario 0: Frequencies of ULI and Person_ID are both missing (equal too)
RFD4682_3$ID_clean<-rep(NA, 214780)
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI==0 &
                   RFD4682_3$Freq_PER==0  ]<-0

# Scenario 1: Frequencies of ULI and Person_ID are equal
length(unique(RFD4682_3$ULI_IDe[(RFD4682_3$Freq_ULI==RFD4682_3$Freq_PER) &
                                 RFD4682_3$Freq_ULI>0                    &
                                 RFD4682_3$Freq_PER>0])
) # N=179,460 This is the majority of cases in the dataset, which is good
# This number matches with the numbers from the confusion tab

RFD4682_3$ID_clean[(RFD4682_3$Freq_ULI==RFD4682_3$Freq_PER) &
                                 RFD4682_3$Freq_ULI>0       &
                                 RFD4682_3$Freq_PER>0]<-1

# Scenario 2: only ULI, Person_ID are missing
# Number estimated using confusion tab: 107, 108 records
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI>0 &
                   RFD4682_3$Freq_PER==0  ]<-2

# Scenario 3, ULI missing but they have PERSON_ID
# Number estimated using confusion tab: 7904, 8433 records
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI==0 &
                   RFD4682_3$Freq_PER>0  ]<-3

# Scenario 4, one PERSON_ID matches with multiple ULI_IDs
# Number estimated using confusion tab: 6, 10 records
RFD4682_3$ID_clean[RFD4682_3$Freq_ULI==1 &
                   RFD4682_3$Freq_PER==2  ]<-4 # this would only fix 6 records

# Fix the rest 4 manually
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="9a4299b85800" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="aa6526b42bfe" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="e054534f4352" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4
RFD4682_3$ID_clean[RFD4682_3$PERSON_IDe=="fcb0ca36fbb4" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-4

RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="9a4299b85800" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"3c32a58470ee228"
RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="aa6526b42bfe" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"f21a83130a09b98"
RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="e054534f4352" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"c80b7fbde166521"
RFD4682_3$ULI_IDe[RFD4682_3$PERSON_IDe=="fcb0ca36fbb4" &
                   is.na(RFD4682_3$ULI_IDe)==TRUE]<-"08bc00a9a5e8720"

# The rest are scenario 5
# N records = 2775, number matches those on confusion cross-tab
RFD4682_3$ID_clean[is.na(RFD4682_3$ID_clean)==TRUE]<-5

# Rename the variable as "scenario"
RFD4682_3<-rename(RFD4682_3, Scenario = ID_clean)
table(RFD4682_3$Scenario, useNA = 'ifany')

# Apply approperiate individual ids to each records, according to its scenario
RFD4682_3$clean_IDe<-rep(NA, 214780)
RFD4682_3$clean_IDe<-if_else(RFD4682_3$Scenario %in% c(3), RFD4682_3$PERSON_IDe, RFD4682_3$ULI_IDe)
```

#### 1.2 Scenario Examples Showcase

```{r}
# Example with each Scenarios
RFD4682_3[RFD4682_3$Freq_ULI==6 & RFD4682_3$Freq_PER==5 ,]     # Scenario 5
RFD4682_3[RFD4682_3$Freq_ULI==15 & RFD4682_3$Freq_PER==15 ,]   # Scenario 1
RFD4682_3[RFD4682_3$Freq_ULI==2 & RFD4682_3$Freq_PER==0 ,]     # Scenario 2
RFD4682_3[RFD4682_3$Freq_ULI==0 & RFD4682_3$Freq_PER==5 ,]     # Scenario 3
RFD4682_3[RFD4682_3$Freq_ULI==1 & RFD4682_3$Freq_PER==2 ,]     # Scenario 4
RFD4682_3[RFD4682_3$Freq_ULI==0 & RFD4682_3$Freq_PER==0 ,]     # Scenario 0
```

#### 1.3 Data cleaning

##### Add clean_IDe and Scenario back to dataset (encrypted)

```{r}
RFD4682_e<-merge(RFD4682_e, select(RFD4682_3, order_ID, Scenario, clean_IDe),
                      by = 'order_ID', all = TRUE)
```

##### ULI ID and PERSON_ID, in NAAT+ subset

```{r}
## ULI_ID: Unique Lifetime Identifier (ULI), numerical values of Alberta PHN.
length(unique(RFD4682_NAAT$ULI_ID))     # N = 5933
table(is.na(RFD4682_NAAT$ULI_ID))       # NA = 0

## PERSON_ID: assigned from the originating LIS
length(unique(RFD4682_NAAT$PERSON_ID)) # N = 5940
table(is.na(RFD4682_NAAT$PERSON_ID))   # NA = 0

## Check the discrepancies between those two
tab_ULI_ID<-table(RFD4682_NAAT$ULI_ID)
tab_ULI_ID.dat<-as.data.frame(tab_ULI_ID)
colnames(tab_ULI_ID.dat)<-c('ULI_ID','Freq_ULI')
tab_ULI_ID.dat.2<-tab_ULI_ID.dat[tab_ULI_ID.dat$Freq>1,]

tab_PERSON_ID<-table(RFD4682_NAAT$PERSON_ID)
tab_PERSON_ID.dat<-as.data.frame(tab_PERSON_ID)
colnames(tab_PERSON_ID.dat)<-c('PERSON_ID','Freq_PER')
tab_PERSON_ID.dat.2<-tab_PERSON_ID.dat[tab_PERSON_ID.dat$Freq>1,]

## Add freq var to 
RFD4682_NAAT_3<-select(RFD4682_NAAT, SERO_ID, ULI_ID, PERSON_ID)
RFD4682_NAAT_3<-merge(RFD4682_NAAT_3, tab_ULI_ID.dat.2, by='ULI_ID',
      all.x = T, all.y = F)
RFD4682_NAAT_3<-merge(RFD4682_NAAT_3, tab_PERSON_ID.dat.2, by='PERSON_ID',
      all.x = T, all.y = F)

RFD4682_NAAT_4<-RFD4682_NAAT_3[is.na(RFD4682_NAAT_3$Freq_ULI)==F |
               is.na(RFD4682_NAAT_3$Freq_PER)==F,]

RFD4682_NAAT_4$freq_minus<-RFD4682_NAAT_4$Freq_ULI-RFD4682_NAAT_4$Freq_PER

print(RFD4682_NAAT_4[RFD4682_NAAT_4$freq_minus>0 |
               is.na(RFD4682_NAAT_4$freq_minus)==TRUE,])

## Save the dataset
save(RFD4682_e, file = '~/APL/RFD4682e.RData')

## Conclusion:
## Use ULI_ID, in NAAT+ subset, seven people have 2 diff PERSON_IDs
## Unique ULI_ID: N=5940
```

```{r}
length(unique(RFD4682_e$clean_IDe))
table(is.na(unique(RFD4682_e$clean_IDe)), useNA = 'ifany')

# Unique Individual in the dataset: 188,722 (beware of missing values)
```

##### Check N=2865 cases with missing in both ULI and Person_ID

```{r}
RFD4682_e[RFD4682_e$Scenario==0,]
## Surprisingly, most of the records in this group has some useful variables,
## such as: age, gender and Anti_RBD test results

# Age
table(RFD4682_e$AGE_AT_COLLECTION[RFD4682_e$Scenario==0], useNA = 'ifany') # 0.03%
# Gender
table(RFD4682_e$GENDER[RFD4682_e$Scenario==0], useNA = 'ifany') # 0.03%
# N-Ig
table(RFD4682_e[RFD4682_e$Scenario==0,]$`N-IgG_RESULT`, useNA = 'ifany') # Can't show all
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$`N-IgG_RESULT`), useNA = 'ifany') # 92.95%
table(is.na(RFD4682_e$`N-IgG_RESULT`), useNA = 'ifany') # 92.95%

# Ig Anti-RBD
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$`RBD-IgGII_RESULT`), useNA = 'ifany')
# no missing, good

# FSA
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$PAT_FSA), useNA = 'ifany') # 9.49%

# Vaccination
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$VACCINATION_STATUS), useNA = 'ifany')
# No miss, good
table(RFD4682_e[RFD4682_e$Scenario==0,]$VACCINATION_STATUS, useNA = 'ifany')
table(RFD4682_e$VACCINATION_STATUS, useNA = 'ifany')

# NAAT Status
table(is.na(RFD4682_e[RFD4682_e$Scenario==0,]$NAAT_STATUS), useNA = 'ifany') # 92.98%
table(is.na(RFD4682_e$NAAT_STATUS), useNA = 'ifany')

id_temp<-RFD4682_e$SERO_ID[RFD4682_e$VACCINATION_STATUS %in% c("Pre-Immunization")]
View(id_temp)
max(id_temp, na.rm = T) # 2020 Nov
min(id_temp, na.rm = T)

# Conclusion: The data quality of this part is good, but we cannot use those in our project
```

##### The Whole APL dataset

```{r}
# The NAAT status variable
table(RFD4682_e$NAAT_STATUS, useNA = 'ifany')

# The Vaccination Status Variable
table(RFD4682_e$VACCINATION_STATUS, useNA = 'ifany')

# Cross-tab of those two
table(RFD4682_e$NAAT_STATUS, RFD4682_e$VACCINATION_STATUS, useNA = 'ifany')
```

##### Categorize the time since last NAAT test

```{r}
hist(RFD4682_e$DAYS_SINCE_NAAT, breaks=20)

# Create an indicator
RFD4682_e$DAYS_NAAT_cat<-rep(NA, 214780)
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>=0 &
                        RFD4682_e$DAYS_SINCE_NAAT<=90 ]<-1  # <3 months
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>90 &
                        RFD4682_e$DAYS_SINCE_NAAT<=180 ]<-2  # 4-6 months
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>180 &
                        RFD4682_e$DAYS_SINCE_NAAT<=270 ]<-3  # 4-6 months
RFD4682_e$DAYS_NAAT_cat[RFD4682_e$DAYS_SINCE_NAAT>270 &
                        is.na(RFD4682_e$DAYS_SINCE_NAAT)==F ]<-4  # 4-6 months
table(RFD4682_e$DAYS_NAAT_cat, useNA = 'ifany')

# Cross-tabulation between NAAT and NAAT_DAYS categories
# Anti-N groups
table(RFD4682_e$DAYS_NAAT_cat[is.na(RFD4682_e$`N-IgG_RESULT`)==F], 
      RFD4682_e$NAAT_STATUS[is.na(RFD4682_e$`N-IgG_RESULT`)==F], useNA = 'ifany')

# Anti-RBD groups
table(RFD4682_e$DAYS_NAAT_cat[is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F &
                              RFD4682_e$VACCINATION_STATUS %in% c('Not Immunized')], 
      RFD4682_e$NAAT_STATUS[is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F &
                              RFD4682_e$VACCINATION_STATUS %in% c('Not Immunized')], useNA = 'ifany')
```

##### Vaccination Status

```{r}
table(RFD4682_NAAT$VACCINATION_STATUS, useNA = 'ifany')
hist(RFD4682_NAAT$DAYS_SINCE_IMM)

# Create day gap

## NAs in NAAT Days var
table(is.na(RFD4682_NAAT$DAYS_SINCE_NAAT)) # Great, no NAs

## NAs in Vaccination Days var
table(is.na(RFD4682_NAAT$DAYS_SINCE_IMM)) # Only 911 missing, not match with Vacc Status

## Calculate the Gap
RFD4682_NAAT$DAYS_Gap<-rep(NA,6165)
RFD4682_NAAT$DAYS_Gap<-RFD4682_NAAT$DAYS_SINCE_NAAT - RFD4682_NAAT$DAYS_SINCE_IMM
table(is.na(RFD4682_NAAT$DAYS_Gap)) #911 Missing, 0 vaccinated at the same day of sample collection
hist(RFD4682_NAAT$DAYS_Gap, 
     main = 'Gap between NAAT and Vaccination',
     xlab = 'Days')

## Create the indicator
RFD4682_NAAT$Vacc_status<-rep(NA, 6165)
RFD4682_NAAT$Vacc_status[RFD4682_NAAT$VACCINATION_STATUS %in% c("Not Immunized")]<-1
RFD4682_NAAT$Vacc_status[RFD4682_NAAT$DAYS_Gap>=0 & is.na(RFD4682_NAAT$DAYS_Gap)==F]<-2
RFD4682_NAAT$Vacc_status[RFD4682_NAAT$DAYS_Gap<0 & is.na(RFD4682_NAAT$DAYS_Gap)==F]<-3
table(RFD4682_NAAT$Vacc_status, useNA = 'ifany')
```

Estimation on number of individuals, using cleaned version of IDe
```{r}
# Total num of individual in the data:
table(is.na(unique(RFD4682_e$clean_IDe)), useNA = 'ifany')    # N=188,722, plus 1 NA

# Number of participants with available Anti-N results:
antin_na<-is.na(unique(RFD4682_e[is.na(RFD4682_e$clean_IDe)==F &
                             is.na(RFD4682_e$`N-IgG_RESULT`)==F,
                             ]$clean_IDe))
table(antin_na, useNA = 'ifany')  # N=39450

# Number of participants with available Anti-S results:
antis_na<-is.na(unique(RFD4682_e[is.na(RFD4682_e$clean_IDe)==F &
                             is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F,
                             ]$clean_IDe))
table(antin_na, useNA = 'ifany') # N=188,688

table(is.na(RFD4682_e$`N-IgG_RESULT`)==F,
      is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F, useNA = 'ifany')
```

Some additional cleaning/editing, making anti-S quantitative
```{r}
# Missing values
summary(is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F) 
# N=36, beware of "not available" values, better recode those as NAs

# Remove all the numbers, see what left
S_chr<-RFD4682_e$`RBD-IgGII_RESULT` %>%
  gsub("[[:digit:]]+", "",.)
table(S_chr, useNA = 'ifany')
## As we can see, the use of units in this data is quite messy, needs additonal cleaning

# How many of them use unit "AU/ml" (arbitury unit)?
AU<-RFD4682_e$`RBD-IgGII_RESULT` %>%
  grepl("ml", .,fixed = F, ignore.case = T)
table(AU) # N=109798

# index
index<-RFD4682_e$`RBD-IgGII_RESULT` %>%
  grepl("index", .,fixed = F, ignore.case = T)
table(index) # N=104946
```

#### 1.4 Other data cleaning

##### Yuan found this: anti-N is only available for the Omicron stage
**Notes:** This is an error in two ver of APL. The anti-N assay results are available and misplaced in the anti-S-RBD column.
```{r}
# The APL Data:
RFD4682_e$COLLECTION_DATE %>%
  summary()

# Anti-N assay results is only available to N=39,450 participants
RFD4682_e[is.na(RFD4682_e$clean_IDe)==F &
          is.na(RFD4682_e$`N-IgG_RESULT`)==F,]$COLLECTION_DATE %>%
  summary()
# Anti-N: 2021/12/14 -- 2022/10/14

# Anti-S assay results is available to most of the participants
RFD4682_e[is.na(RFD4682_e$clean_IDe)==F &
          is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F,]$COLLECTION_DATE %>%
  summary()
```

#### As required by Matthew, add quintmat & qunitsoc variable

```{r}
# load a matching scale, by 6-digit postal codes
pcode_match <- read.delim("~/COVID-assay-adjust-methods/1_data/private/ot_ccp_postalcodecodeequivalence.txt")

# load updated version of APL

View(RFD4682_v2)

# match 6-digit postal codes to the first ver
RFD4682_e$PAT_POSTALCODE <- RFD4682_v2$PAT_POSTALCODE

# cleanup NA
length(RFD4682_e$PAT_POSTALCODE[RFD4682_e$PAT_POSTALCODE=="N/A"]) # needs to cleanup "N/A"
RFD4682_e$PAT_POSTALCODE[RFD4682_e$PAT_POSTALCODE=="N/A"]<-NA
table(is.na(RFD4682_e$PAT_POSTALCODE))

# does the postal code match?
# create a truncated ver of 6-digit postal code
RFD4682_e$PAT3<-substr(RFD4682_e$PAT_POSTALCODE, 1, 3)

RFD4682_e$PAT_match<-if_else(RFD4682_e$PAT_FSA==RFD4682_e$PAT3, 1, 0)
table(RFD4682_e$PAT_match, useNA = "ifany") # 1065 unmatch, 3276 missing

RFD4682_e %>%
  .[.$PAT_match == 0 &  !is.na(.$PAT3),] %>%
  select(PAT_POSTALCODE, PAT3, PAT_FSA)
# N=1756 unmatched cases btw APL ver 1 & 2. Talked to M. Knight and he would like to use the updated ver

RFD4682_e<-merge(RFD4682_e, select(pcode_match, PCODE, QUINTMAT, QUINTSOC),
      by.x = "PAT_POSTALCODE", by.y="PCODE", all.x = T, all.y = F)

# a clean version of if for M.Knight
RFD4682_e1<-select(RFD4682_e, -c("PAT3", "PAT_match", "N", "Col_Month"))

```

#### A quick clean of the Anti-N data

```{r}
# Remove unecessary 'index'
table(RFD4682_e$`N-IgG_RESULT`)

RFD4682_e$N<-RFD4682_e$`N-IgG_RESULT` %>%
  gsub("Index", "",.) %>%
  as.numeric()

hist(RFD4682_e$N, breaks=25)

# Results
table(RFD4682_e$`N-IgG_INTERP`, useNA = 'ifany')
```

#### Add a year-month var

```{r}

# only keep the month, remove the data
RFD4682_e$Col_Month<-format(as.Date(RFD4682_e$COLLECTION_DATE), "%Y-%m")
table(RFD4682_e$Col_Month, useNA='ifany')

# as char variable if needed for calculation, use as.yearmon() from {zoo} or just use the Col_Date var
```

#### 1.5 Save a cleaned ver of RFD4682e

```{r}
save(RFD4682_e, file = "../1_data/private/RFD4682e.RData")

save(RFD4682_e1, file = "../1_data/private/RFD4682e1.RData") # ver for matt only
```

## 2. APL v2 (Jan 2024) - Data Cleaning

#### ADD order ID and other ids -- doesn't work
```{r}
RFD4682_v2$order_ID<-RFD4682_e$order_ID

cbind.data.frame(RFD4682_e$order_ID, RFD4682_e$SERO_ID, 
                 RFD4682_v2$order_ID, RFD4682_v2$SERO_ID)
```

#### 2.1 De-duplicate
```{r}
# Does duplicated records exist?
table(duplicated(RFD4682_v2))  # no-all records are unique

# use the saved duplicated SERO_ID variables to de_duplicate:
RFD4682_v2dup<-RFD4682_v2 %>%
 .[duplicated(RFD4682_v2$SERO_ID)==T,]

# Duplicated records to clean: 
RFD4682_v2dup2<-RFD4682_v2 %>%
  .[.$SERO_ID %in% RFD4682_v2dup$SERO_ID,]
RFD4682_v2dup2$SERO_ID[9:16]<-paste0(RFD4682_v2dup2$SERO_ID[9:16],
                                     rep(c("_1", "_2"),4))

# Clean, now SERO_ID is unique
RFD4682_v2<-RFD4682_v2 %>%
  .[!(.$SERO_ID %in% RFD4682_v2dup$SERO_ID),]

RFD4682_v2<-rbind.data.frame(RFD4682_v2, RFD4682_v2dup2[c(2,4,6,8,9:16),])
rm(RFD4682_v2dup, RFD4682_v2dup2)
```
##### De-duplicate SERO_ID for APL v1
```{r}
# use the saved duplicated SERO_ID variables to de_duplicate:
RFD4682_edup<-RFD4682_e %>%
 .[duplicated(RFD4682_e$SERO_ID)==T,]

# Duplicated records to clean: 
RFD4682_edup2<-RFD4682_e %>%
  .[.$SERO_ID %in% RFD4682_edup$SERO_ID,]
RFD4682_edup2<-RFD4682_edup2 %>%
  arrange(order_ID)
RFD4682_edup2$SERO_ID[9:16]<-paste0(RFD4682_edup2$SERO_ID[9:16],
                                     rep(c("_1", "_2"),4))

# Clean, now SERO_ID is unique
RFD4682_e<-RFD4682_e %>%
  .[!(.$SERO_ID %in% RFD4682_edup$SERO_ID),]

RFD4682_e<-rbind.data.frame(RFD4682_e, RFD4682_edup2[c(2,4,6,8,9:16),])
rm(RFD4682_edup, RFD4682_edup2)
```

#### 2.2 Add ULI_IDe from v1
```{r}
library(duawranglr)    # use this to encrypt the ids

## get crosswalk and admin data files
dua_cw_file <- system.file('extdata', 'dua_cw.csv', package = 'duawranglr')
admin_file <- system.file('extdata', 'admin_data.csv', package = 'duawranglr')

## set the DUA crosswalk
set_dua_cw(dua_cw_file)
```

```{r}
# Use SERO_ID and 'RBD-IgGII_RESULT' to identify each records
RFD4682_v2<-merge(RFD4682_v2, select(RFD4682_e, SERO_ID, order_ID, PERSON_IDe, ULI_IDe,
                         clean_IDe, QUINTMAT, QUINTSOC), 
      by ="SERO_ID", all = T
      )
# ULI - ULI_IDe crosswalk
ULI_crosswalk<-select(RFD4682_v2, ULI_ID, ULI_IDe)
ULI_crosswalk<-distinct(ULI_crosswalk)

table(is.na(ULI_crosswalk$ULI_IDe))
ULI_crosswalk %>%
  .[is.na(.$ULI_IDe),]

# APL ver2 seems have solved missing value issue in the ULI ID, needs to add those
table(is.na(RFD4682_v2$ULI_ID), is.na(RFD4682_v2$ULI_IDe))

table(is.na(RFD4682_v2$PERSON_ID), is.na(RFD4682_v2$PERSON_IDe))

# encrypt new ULI_ID
set.seed(240611)
ULI_crosswalk_e <- deid_dua(ULI_crosswalk, 
                      id_col = 'ULI_ID',
                      new_id_name = 'ULI_IDe2',
                      write_crosswalk = F, id_length = 15)

ULI_crosswalk$ULI_IDe2<-ULI_crosswalk_e$ULI_IDe2

# clean the ULI_ID crosswalk file
table(is.na(ULI_crosswalk$ULI_ID))
ULI_crosswalk<-ULI_crosswalk[!is.na(ULI_crosswalk$ULI_ID),]
ULI_crosswalk<-ULI_crosswalk[-64591,]
ULI_crosswalk$ULI_IDe[is.na(ULI_crosswalk$ULI_IDe)]<-ULI_crosswalk$ULI_IDe2[is.na(ULI_crosswalk$ULI_IDe)]

# apply the cross-walk to the v2 data
df.temp1<-RFD4682_v2[!is.na(RFD4682_v2$ULI_ID) & is.na(RFD4682_v2$clean_IDe),]
df.temp2<-RFD4682_v2[!(!is.na(RFD4682_v2$ULI_ID) & is.na(RFD4682_v2$clean_IDe)),]

df.temp1<-merge(select(df.temp1, -c("ULI_IDe")), 
             ULI_crosswalk[,1:2],
      by="ULI_ID", all.x = T, all.y = F)
df.temp1$clean_IDe<-df.temp1$ULI_IDe

RFD4682_v2<-bind_rows(df.temp1, df.temp2)
rm(df.temp1, df.temp2, ULI_crosswalk_e)
```

```{r}
# Still, there's some missing value in clean_IDe
RFD4682_v2 %>%
  .[is.na(.$clean_IDe),]
```
#### 2.3 Clean anti-N and anti-S

##### 2.3.1 Anti-N cleaning (remove index)
```{r}
# Create an indicator for misplaced anti-N
RFD4682_v2$abb_s_index<-grepl("index", RFD4682_v2$`RBD-IgGII_RESULT`, ignore.case = T)
RFD4682_v2$abb_s_AU<-grepl("ml", RFD4682_v2$`RBD-IgGII_RESULT`, ignore.case = T)
table(RFD4682_v2$abb_s_index, RFD4682_v2$abb_s_AU, useNA = 'ifany')
# 36 samples missing abbott_s info

RFD4682_v2$misplace<-rep("anti_S missing", 214776)
RFD4682_v2$misplace[RFD4682_v2$abb_s_index==T]<-"index"
RFD4682_v2$misplace[RFD4682_v2$abb_s_AU==T]<-"AU"
table(RFD4682_v2$misplace, RFD4682_v2$abb_s_index, useNA = 'ifany')

RFD4682_v2$`N-IgG_RESULT`[RFD4682_v2$misplace=="index"]<-
  RFD4682_v2$`RBD-IgGII_RESULT`[RFD4682_v2$misplace=="index"]

RFD4682_v2$`RBD-IgGII_RESULT`[RFD4682_v2$misplace=="index"]<-NA

RFD4682_v2<-select(RFD4682_v2, -c("abb_s_index", "abb_s_AU"))
```

```{r}
# Remove "index" for anti-N
RFD4682_v2$`N-IgG_RESULT`<-gsub('Index', '', RFD4682_v2$`N-IgG_RESULT`)
RFD4682_v2$`N-IgG_RESULT`<-as.numeric(RFD4682_v2$`N-IgG_RESULT`)

# Anti-N interp:
summary(RFD4682_v2$`N-IgG_RESULT`[RFD4682_v2$`N-IgG_INTERP`=="Positive"]) # good
summary(RFD4682_v2$`N-IgG_RESULT`[RFD4682_v2$`N-IgG_INTERP`=="Negative"]) # good
summary(RFD4682_v2$`N-IgG_RESULT`[is.na(RFD4682_v2$`N-IgG_INTERP`)]) # needs correction

RFD4682_v2$`N-IgG_INTERP`[RFD4682_v2$`N-IgG_RESULT` < 1.4 & 
                          !is.na(RFD4682_v2$`N-IgG_RESULT`)]<-"Negative"
RFD4682_v2$`N-IgG_INTERP`[RFD4682_v2$`N-IgG_RESULT` >= 1.4 & 
                          !is.na(RFD4682_v2$`N-IgG_RESULT`)]<-"Positive"

summary(RFD4682_v2$COLLECTION_DATE[!is.na(RFD4682_v2$`N-IgG_INTERP`)])
```

##### 2.3.2 Anti-S cleaning (remove AU/ml)
```{r}
RFD4682_v2$`RBD-IgGII_RESULT`<-gsub('[a-z]', '', RFD4682_v2$`RBD-IgGII_RESULT`)
RFD4682_v2$`RBD-IgGII_RESULT`<-gsub('[A-Z]', '', RFD4682_v2$`RBD-IgGII_RESULT`)
RFD4682_v2$`RBD-IgGII_RESULT`<-gsub('[/<>]', '', RFD4682_v2$`RBD-IgGII_RESULT`)
RFD4682_v2$`RBD-IgGII_RESULT`<-as.numeric(RFD4682_v2$`RBD-IgGII_RESULT`)
summary(RFD4682_v2$`RBD-IgGII_RESULT`)

# >800000 is recoded as 800000
RFD4682_v2 %>%
  .[.$`RBD-IgGII_RESULT`<800000 & 100000<.$`RBD-IgGII_RESULT` &
    !is.na(.$`RBD-IgGII_RESULT`),]

# Double check in anti_S interpretation
summary(RFD4682_v2$`RBD-IgGII_RESULT`[RFD4682_v2$`RBD-IgGII_INTERP`=="Positive"]) # -->NA
summary(RFD4682_v2$`RBD-IgGII_RESULT`[RFD4682_v2$`RBD-IgGII_INTERP`=="Negative"]) # -->NA
summary(RFD4682_v2$`RBD-IgGII_RESULT`[is.na(RFD4682_v2$`RBD-IgGII_INTERP`)]) # good

# correct anti-S interpretations
RFD4682_v2$`RBD-IgGII_INTERP`[is.na(RFD4682_v2$`RBD-IgGII_RESULT`)]<-NA

table(RFD4682_v2$`RBD-IgGII_INTERP`, useNA = 'ifany')

```
#### 2.4 Remove unencryoted individual inditifier
```{r}
RFD4682_ev2<-select(RFD4682_v2, -c(ULI_ID, PERSON_ID, misplace))
```

#### 2.5 Save the encrypted ver of APL data
```{r}
save(RFD4682_v2, file = "../1_data/private/RFD4682_v2.RData")
save(RFD4682_ev2, file = "../1_data/private/RFD4682_ev2.RData")

save(RFD4682_ev2, file = "~/../../jmy/APL/RFD4682_ev2.RData")
```

## 3. Clean other two APL sub datasets (Vax, NAAT)

```{r}
# Add encrypted IDe to vax and NAAT data
RFD4682_v2_vaxe<-merge(RFD4682_v2_vax, select(ULI_crosswalk, ULI_ID, ULI_IDe),
      by.x="ULI", by.y="ULI_ID", all.x = T, all.y = F)
# any missing?
table(is.na(RFD4682_v2_vaxe$ULI_IDe))

# also encrypt the ULI_ID in the RFD4682_vax

list_temp<-RFD4682_v2_vaxe %>%
  .[is.na(.$ULI_IDe),]
```